### 2011-05-12 report/report.tex
Add report template

%
% $Id$ 
%
% $LastChangedDate$ 
% 
% $LastChangedBy$
%

\documentclass[10pt]{report}
\usepackage{graphicx}
\usepackage{setspace}			
\onehalfspacing
% numbering down to subsubsection
\setcounter{secnumdepth}{3}

\title{CS6999 Web Data and Text Mining:\\Parallel Text Indexer Project Report}
\author{Damien DuBios\\Justin Kamerman 3335272\\Raman Singh}
\date{\today}

\begin{document}
\maketitle
% No chapter numbers
\renewcommand*\thesection{\arabic{section}}

%----------------------------------------
% Introduction
%----------------------------------------
\section{Introduction}
Briefly in at most one page describe the main idea of the project and why is it important\\

%----------------------------------------
% Literature Review
%----------------------------------------
\section{Literature Review}
\label{sec:literaturereview}
The Aho-Corasick string matching algorithm\cite{RefWorks:103} is a
kind of dictionary matching search algorithm that constructs a finite
state machine to scan for a given set of keywords. It is, in effect, a
reduced grammar regular expression parser described in
\cite{RefWorks:111}. In our search implementation, the finite state
machine is constructed from a list of keywords. Then the documents
of the test corpus are read from secondary storage and fed through the
finite state machine. As soon as a document is found to contain at
least one instance of each of a set of search terms, parsing of that document
ends.

%----------------------------------------
% Motivation
%----------------------------------------
\section{Motivation}
\label{sec:motivation}

%----------------------------------------
% Research Plan
%----------------------------------------
\section{Research Plan}
\label{sec:researchplan}

%----------------------------------------
% Design Component
%----------------------------------------
\section{Design Component}
\label{sec:designcomponent}
A black-box functional view of the indexing system is show in figure
\ref{fig:blackbox}. The system proceses a continuous incoming flow of
documents and distribute them to parallel indexing threads running on
physically seprate, heterogenous nodes.  Document queries are performed
using an evolving search index. The operational balance is to
timeously index new additions to the corpus while servicing concurrent
search requests. A view as to how the components of the indexer system
are deployed is show in figure \ref{fig:deploymentmodel}.

As can be seen in figure \ref{fig:deploymentmodel}, the indexer
processes are symmetrical and deployed on muliple physical nodes. The
individual indexers do not interact directly with one another, making
for a simple deployment and operation model. The execution loop of
each indexer is as follows:

\begin{enumerate}
\item Initialize the indexer by retrieving a collection of index
  keywords and synonyms from a relational databse. These keywords are
  used to construct a lexical parser which will be used to scan and
  index documents.
\item Retrieve a batch of unprocessed documents from a relational
  database. The document batch will be sized according to the physical
  capabilities of each node. In this implementation, this tuning task
  is a manual exercise but future enhancements may include an adaptive
  loading component.
\item Parse each document retrieved and contruct an inverted index
  representing the batch. This \textit{delta} index, as we shall call
  it, is then used to augment the global index maintained in a
  relational database.
\item Repeat from step 2.
\end{enumerate}

As shown in figure \ref{fig:deploymentmodel}, the searcher component
can be executed from any node which has access to the relational
database housing the document index. The execution path of a single
search query wold be as follows:

\begin{enumerate}
  \item Canonize search terms based on keyword synonyms defined in the
    keyword store.
  \item Execute a boolean query against the inverted index.
  \item Return the list of corpus documents containing the
    intersection of the canons of the search terms.
\end{enumerate}

\begin{figure}
  \begin{center}
	\includegraphics[width=\textwidth,height=!]{blackbox}
  \end{center}
  \caption{A black-box view of the indexer}
  \label{fig:blackbox}
\end{figure}

\begin{figure}
  \begin{center}
	\includegraphics[width=\textwidth,height=!]{deploymentmodel}
  \end{center}
  \caption{Indexer deployment model}
  \label{fig:deploymentmodel}
\end{figure}

A guiding architectural principal of the indexer design is to separate the implementation
into three logical layers:
\begin{itemize}
\item \textbf{Software Layer:} implements parsing, indexing and search
  algorithms.
\item \textbf{Data Access Layer:} implements an
  \textit{object-relational mapping}, insulating algorithm
  implementation from the persistence details. This layer implements
  classes which encapsulate persistent entities within the system.
\item \textbf{Database:} the persistent store for the document
  collection and inverted index.
\end{itemize}

The implementation classes and their relationships are represented in
a UML class diagram in figure \ref{fig:overallclassdiagram}. Figure
\ref{fig:overallsequencediagram} is a UML sequence diagram showing how
the layers interact at a high level.  The implementation of each of
these layers is described in more detail in section
\ref{sec:implementation}.

\begin{figure}
  \begin{center}
	\includegraphics[width=\textwidth,height=!]{overallclassdiagram}
  \end{center}
  \caption{Overall UML class diagram design}
  \label{fig:overallclassdiagram}
\end{figure}

\begin{figure}
  \begin{center}
	\includegraphics[width=\textwidth,height=!]{overallsequencediagram}
  \end{center}
  \caption{Overall UML sequence diagram design}
  \label{fig:overallsequencediagram}
\end{figure}

%----------------------------------------
% Implementation 
%----------------------------------------
\section{Implementation}
\label{sec:implementation}
The individual worker processes of the indexer system are implemented as Java
programs. The programs make use of the following thirdparty libraries:

\begin{itemize}
  \item \textbf{DBPool:} a database connection pooling library, used
    to manage connections to relational databases.
  \item \textbf{Apache Commons Logging:} a generic logging API used by
    DBPool.
  \item \textbf{Apache Commons CLI:} a library for parsing command
    line arguments passed to Java programs.
  \item \textbf{MySQL JDBC driver:} the MySQL Java client driver.
\end{itemize}

The indexer system maintains an \textit{inverted index} of an evolving
corpus and services concurrent search requests against the index. The
current implementation is based on boolean retrieval methods described
in \cite{RefWorks:109}. An Aho-Corasick state machine
\cite{RefWorks:103} is used to scan documents for a predefined set of
keywords, and construct a term index for each document (delta
index). The term occurrences for each document are used to augment an
existing index for documents already processed. The document
collection, index, and keyword list are maintained in a single MySQL
database instance. It is through this database that the parallel
indexer instances are initialized and synchronized. This configuraion
is operationally simple and easy to implement but has the potential to
become a throttle point as the system is scaled to larger and
faster-evolving document collections, and is required to service a
higher rate of queries. In order to scale the current implementation
to large document collections, a more efficient mechanism would be
required in this repect.

The following subsections detail the implementation of the
architectural layers identified in section \ref{sec:designcomponent}.

\subsection{Data Access Layer}
\label{sec:dataaccesslayer}
Access to persistent storage is via objects of the data access
layer. Confining data access to a logical grouping of classes
insulates the other application components from changes in database
schema.

\subsubsection{Document}
\label{sec:document}
The DocumentDAO singleton class provides methods for retrieving
Document objects from the document database: List <Document>
getDocuments (int count): retrieve a specified number of unprocessed
documents from the database. The documents retrieved will be
atomically marked as processed so that processing is not duplicated on
other nodes. In this scheme, a document is considered processed to all
other processes once retrieved from the database. This may cause
problems from a fault recovery perspective if the processing node
fails. A proposed enhancement to this scheme is to use another state,
processing, to indicate that the document has been retrieved and to
update the state to a processed state once processing is actually
complete. For the initial implementation we will use the processed
state only.
Document getDocumentById (int id): retrieve a specific document from the database, referenced by its ID.
Int getID (): accessor method.
String getContent (): accessor method.
String getAuthor (): accessor method.
String getTitle (): accessor method
String getLink(): accessor method.

The Document class encapsulates the BLOG database table, with the
addition of a state field that tracks whether a document has been
processed or not. This field will be implemented as a new field on the
BLOG table or as a new reference table DOCUMENT_STATE.  and show UML
class and sequence diagrams respectively for non-trivial classes and
methods involved in the implementation of the Document data access
layer.

%----------------------------------------
% Experiment Results
%----------------------------------------
\section{Experiment Results}
\label{sec:experimentresults}
FIX: All tests were run on a Dell laptop with an Intel Core Duo 2GHz
processor, 1GB RAM, running a 32 bit Linux 2.6.37 kernel. The Java
Virtual Machine used was version 1.6.0-24.

\subsection{Aho-Corasick}
Various tests were conducted to characterize the performance of the
Aho-Corasick search algorithm:

\begin{itemize}
\item The time taken to construct the Aho-Corasick state machine was
  measured for different numbers of keywords. The results of this test
  is shown in figure \ref{fig:ahostatemachine} and construction time
  can easily be considered linear with respect to the number of
  keywords. This is consistent with \cite{RefWorks:103} which proves
  that the state machine construction algorithm is linearly
  proportional to the sum of the lengths of the keywords used to
  construct the state machine.

\item The time taken for the Aho-Corasick state machine to scan
  different sized corpora was measured. The test was repeated for
  various state machines, constructed using a 100, 200, and 400
  keyword set. The results of this test are shown on figure
  \ref{fig:ahoscan} and indicate that scan time is nearly linear with
  respect to the size of the corpus and independent of the number of
  keywords used to generate the state machine. This last result is
  consistent with \cite{RefWorks:103} which proves that the number of
  state transitions involved in processing an input string is
  independent of the number of keywords used to construct the state
  machine.

\item The time taken to search for different numbers of keywords was
  measured over different sized corpora. A set of ten keywords was
  selected randomly and the the corpora searched for ${n \choose k}$
  enumerated combinations thereof to obtain an average for a
  particular keyword set size. The results of this test are shown in
  figure \ref{fig:ahosearch} and indicate that search time is
  independent of the number of terms used in the search and linear
  with respect to the size of the corpus. This first result is
  surprising given that the search process exists as soon as a single
  instance of each search term is found. One would expect the number
  of documents scanned completely to increase as the number of search
  term increases. The small average document size and the fact that
  the search hit rate is low may account for this phenomenon.
\end{itemize}

\begin{figure}
  \begin{center}
	\includegraphics[width=\textwidth,height=!]{ahostatemachine}
  \end{center}
  \caption{Aho-Corasick state machine construction}
  \label{fig:ahostatemachine}
\end{figure}

\begin{figure}
  \begin{center}
	\includegraphics[width=\textwidth,height=!]{ahoscan}
  \end{center}
  \caption{Aho-Corasick state machine scanning}
  \label{fig:ahoscan}
\end{figure}

\begin{figure}
  \begin{center}
	\includegraphics[width=\textwidth,height=!]{ahosearch}
  \end{center}
  \caption{Aho-Corasick state machine search}
  \label{fig:ahosearch}
\end{figure}

\subsection*{Inverted Index}
Various tests were conducted to characterize the performance of the
inverted index search algorithm:

\begin{itemize}
\item The time taken to construct the inverted index was measured for
  different sized corpora. The results of this test is shown in figure
  \ref{fig:invscan} and, as expected, indicate a linear relationship
  between index construction time and the size of the corpora.

\item The time taken to search for different numbers of keywords was
  measured over different sized corpora. A set of ten keywords was
  selected randomly and the the corpora searched for ${n \choose k}$
  enumerated combinations thereof to obtain an average for a
  particular keyword set size. The results of this test are shown in
  figure \ref{fig:invsearch}. If the results for one and ten search
  words are ignored, search times appear to be independent of the
  number of search terms and close to linear with respect to the size
  of the corpus. The unusual results for one and ten keywords warrant
  further investigation.
\end{itemize}

\begin{figure}
  \begin{center}
	\includegraphics[width=\textwidth,height=!]{invscan}
  \end{center}
  \caption{Inverted index creation}
  \label{fig:invscan}
\end{figure}

\begin{figure}
  \begin{center}
	\includegraphics[width=\textwidth,height=!]{invsearch}
  \end{center}
  \caption{Inverted index search}
  \label{fig:invsearch}
\end{figure}

\subsection{Comparison}
Comparing corpus scan times for Aho-Corasick and inverted index
implementations, figures \ref{fig:ahoscan} and \ref{fig:invscan}
respectively, indicates that the inverted index scanner is
significantly faster than the Aho-Corasick state machine. Given that
the grammar supported by the JFlex-generated scanner is more expressive
than that of the Aho-Corasick state machine, one would expect results
to more comparable. Reviewing the source code of the generated scanner
indicates a compact and efficient implementation using packed data
structures to represent the internal state machine. In comparison, the
Aho-Corasick implementation is relatively unoptimized, closely
following the pseudo-code in \cite{RefWorks:103}.

Comparing search times for Aho-Corasick and inverted index
implementations, figures \ref{fig:ahosearch} and \ref{fig:invsearch}
respectively, indicates that the inverted index implementation has an
advantage of several orders of magnitude. The initial cost of creating
the inverted index is significantly larger than the creation of the
Aho-Corasick state machine however this cost is amortized over each
subsequent search. The Aho-Corasick implementation rescans the entire
corpus for each search, as opposed to a scan of the inverted index
alone. This exercise is a classic demonstration of the impetus behind
using indexed search techniques for large corpora.

%----------------------------------------
% Conclusions
%----------------------------------------
\section{Conclusions}
\label{sec:conclusions}

%----------------------------------------
% Bibliography
%----------------------------------------
% Change Bibliography title to 'References'
\renewcommand\refname{References}
\bibliography{bibliography}
\bibliographystyle{IEEEannot}

%--------------------------------------------------
\end{document}

### 2011-05-13 report/report.tex
Finished convertion from Word Doc to Latex; Completed report introduction

@@
 \section{Introduction}
-Briefly in at most one page describe the main idea of the project and why is it important\\
+Fundamental to the real-world application of \textit{Information
+  Retrieval} techniques is the ability to objectively measure the
+quality of a particular algorithm or system and compare it to
+others. So often, information retrieval algorithms lack an analytical
+model that would allow one to make formal statements as to the
+performance boundaries of a particular system. In addition, as with
+machine learning and classification problems, using a scalar value to
+represent the performance of a system as a whole is misleading and
+does not account for the variety of operating conditions under which
+the system may be expected to perform.
+
+In measuring the performance of an information retrieval system, the
+following factors constitute a set of operating conditions which
+qualify the result:
+
+\begin{itemize}
+  \item Accuracy of the retrieval result with respect to the
+    query. Common measures in this respect are \textit{Precision} and
+    \textit{Recall}. Search techniques employing latent semantics of
+    document collections are subtly more difficult to quantify, though
+    measurable nonetheless.
+
+    \item Speed with which queries are answered. This metric is a
+      function of the number of search terms and the size of the
+      corpus.
+
+    \item Size of the document collection and, if not static, the rate
+      at which new documents are being added and/or removed.
+
+    \item The computational cost of the technique. Multiple
+      activities may contribute to this aspect: searching, indexing,
+      preprocessing, and post processing. It is important to consider
+      whether the technique allows indices or intermediate
+      approximations of the document collection to be altered
+      incrementally or whether they must be regenerated to incorporate
+      new documents. Also significant is whether the technique can be
+      parallelized and thereby benefit from concurrency.
+
+    \item The storage requirements of the technique. How do the
+      storage requirements evolve with the size of the collection
+      and/or degree of concurrency employed ?
+\end{itemize}
+
+To drive information retrieval research in one direction or another,
+it would be useful to have at ones disposal a physical implementation
+framework within which one could empirically explore the operating
+parameters of a particular algorithm and/or technique along the lines
+described above. Such a system would allow researchers to investigate
+how the system responds to variation in any one of these operating
+parameters. Does the query speed increase as more computational
+capacity is added ? How do storage requirements vary with the size of
+the document collection ? The aim of this project is to create such an
+information retrieval \textit{test-bed}.
+
+The initial reference implementation of this information retrieval
+test-bed is based on a simple \textit{inverted index} but has been
+architected such that the specific algorithm could easily be changed
+without major rework. This paper presents a survey of existing
+literature in support of our test-bed concept, as well as the specific
+algorithmic techniques embodied in the reference implementation. In
+addition, a detailed architectural overview of the system is
+presented, giving insight into how the system would be modified for
+alternative algorithms. Lastly, an evaluation of our inverted index
+reference implementation is conducted and the results analyzed.
 
@@
 \label{sec:literaturereview}
+
 The Aho-Corasick string matching algorithm\cite{RefWorks:103} is a
@@
 reduced grammar regular expression parser described in
-\cite{RefWorks:111}. In our search implementation, the finite state
-machine is constructed from a list of keywords. Then the documents
-of the test corpus are read from secondary storage and fed through the
-finite state machine. As soon as a document is found to contain at
-least one instance of each of a set of search terms, parsing of that document
-ends.
+\cite{RefWorks:111}.
 
@@
 A black-box functional view of the indexing system is show in figure
-\ref{fig:blackbox}. The system proceses a continuous incoming flow of
+\ref{fig:blackbox}. The system processes a continuous incoming flow of
 documents and distribute them to parallel indexing threads running on
-physically seprate, heterogenous nodes.  Document queries are performed
+physically separate, heterogeneous nodes.  Document queries are performed
 using an evolving search index. The operational balance is to
@@
 As can be seen in figure \ref{fig:deploymentmodel}, the indexer
-processes are symmetrical and deployed on muliple physical nodes. The
+processes are symmetrical and deployed on multiple physical nodes. The
 individual indexers do not interact directly with one another, making
@@
 \item Initialize the indexer by retrieving a collection of index
-  keywords and synonyms from a relational databse. These keywords are
+  keywords and synonyms from a relational database. These keywords are
   used to construct a lexical parser which will be used to scan and
@@
   loading component.
-\item Parse each document retrieved and contruct an inverted index
+\item Parse each document retrieved and construct an inverted index
   representing the batch. This \textit{delta} index, as we shall call
@@
 database instance. It is through this database that the parallel
-indexer instances are initialized and synchronized. This configuraion
-is operationally simple and easy to implement but has the potential to
-become a throttle point as the system is scaled to larger and
-faster-evolving document collections, and is required to service a
+indexer instances are initialized and synchronized. This configuration
+is operationally simple and easy to implement but the database has the
+potential to become a throttle point as the system is scaled to larger
+and faster-evolving document collections, and is required to service a
 higher rate of queries. In order to scale the current implementation
 to large document collections, a more efficient mechanism would be
-required in this repect.
+required in this respect.
 
@@
 
+% Data Access Layer
 \subsection{Data Access Layer}
@@
 insulates the other application components from changes in database
-schema.
+schema. The following subsections detail the individual data entities
+handled by the data access later.
 
+% Document
 \subsubsection{Document}
 \label{sec:document}
-The DocumentDAO singleton class provides methods for retrieving
-Document objects from the document database: List <Document>
-getDocuments (int count): retrieve a specified number of unprocessed
-documents from the database. The documents retrieved will be
-atomically marked as processed so that processing is not duplicated on
-other nodes. In this scheme, a document is considered processed to all
-other processes once retrieved from the database. This may cause
-problems from a fault recovery perspective if the processing node
-fails. A proposed enhancement to this scheme is to use another state,
-processing, to indicate that the document has been retrieved and to
-update the state to a processed state once processing is actually
-complete. For the initial implementation we will use the processed
-state only.
-Document getDocumentById (int id): retrieve a specific document from the database, referenced by its ID.
-Int getID (): accessor method.
-String getContent (): accessor method.
-String getAuthor (): accessor method.
-String getTitle (): accessor method
-String getLink(): accessor method.
+The \texttt{DocumentDAO} singleton class provides methods for retrieving
+\texttt{Document} objects from the document database:
 
-The Document class encapsulates the BLOG database table, with the
+\begin{itemize}
+\item \texttt{List<Document> getDocuments (int count)}: retrieve a
+  specified number of unprocessed documents from the database. The
+  documents retrieved will be atomically marked as processed so that
+  processing is not duplicated on other nodes. In this scheme, a
+  document is considered processed to all other processes once
+  retrieved from the database. This may cause problems from a fault
+  recovery perspective if the processing node fails. A proposed
+  enhancement to this scheme is to use another state, processing, to
+  indicate that the document has been retrieved and to update the
+  state to a processed state once processing is actually complete. For
+  the initial implementation we will use the processed state only.
+
+\item \texttt{Document getDocumentById (int id)}: retrieve a specific
+  document from the database, referenced by its ID.
+
+\item \texttt{int getID ()}: accessor method.
+\item \texttt{String getContent ()}: accessor method.
+\item \texttt{String getAuthor ()}: accessor method.
+\item \texttt{String getTitle ()}: accessor method
+\item \texttt{String getLink()}: accessor method.
+\end{itemize}
+
+The \texttt{Document} class encapsulates the BLOG database table, with the
 addition of a state field that tracks whether a document has been
-processed or not. This field will be implemented as a new field on the
-BLOG table or as a new reference table DOCUMENT_STATE.  and show UML
-class and sequence diagrams respectively for non-trivial classes and
-methods involved in the implementation of the Document data access
-layer.
+processed or not. Figures \ref{fig:documentclassdiagram} and
+\ref{fig:documentsequencediagram} show class and sequence diagrams
+respectively for non-trivial classes and methods involved in the
+implementation of the Document data access layer.
+
+\begin{figure}
+  \begin{center}
+	\includegraphics[width=\textwidth,height=!]{documentclassdiagram}
+  \end{center}
+  \caption{Document class diagram}
+  \label{fig:documentclassdiagram}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+	\includegraphics[width=\textwidth,height=!]{documentsequencediagram}
+  \end{center}
+  \caption{Document sequence diagram}
+  \label{fig:documentsequencediagram}
+\end{figure}
+
+% Keyword
+\subsubsection{Keyword}
+The \texttt{KeywordDAO} singleton class provides methods for
+retrieving \texttt{Keyword} objects from the document
+database. Because the keyword set is relatively static and accessed
+often, the full set of keywords will be loaded from the database and
+cached when the class initializes. The class also provides methods for
+converting search terms to their canonical form.
+
+\begin{itemize}
+\item \texttt{List<String> getKeywords ()}: this method
+  will returns a list of all \texttt{Keyword} objects defined in the
+  keyword database.
+
+\item \texttt{List<Integer> canonize (String query)}: return the list
+  of IDs corresponding to a comma-separated list of query terms.
+
+\item \texttt{Keyword getKeywordById (int id)}: return the
+  \texttt{Keyword} object with the given identifier.
+\end{itemize}
+
+The \texttt{Keyword} field encapsulates a canonical search term and a
+list of its synonyms. Figures \ref{fig:keywordclassdiagram} and
+\ref{fig:keywordsequencediagram} show UML class and sequence diagrams
+respectively for non-trivial classes and methods involved in the
+implementation of the Keyword data access layer.
+
+\begin{figure}
+  \begin{center}
+	\includegraphics[width=\textwidth,height=!]{keywordclassdiagram}
+  \end{center}
+  \caption{Keyword class diagram}
+  \label{fig:keywordclassdiagram}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+	\includegraphics[width=\textwidth,height=!]{keywordsequencediagram}
+  \end{center}
+  \caption{Keyword sequence diagram}
+  \label{fig:keywordsequencediagram}
+\end{figure}
+
+% Index
+\subsubsection{Index}
+The \texttt{IndexDAO} class provides methods for updating and
+searching the inverted index table stored in the database. There are
+two methods defined in this class:
+
+\begin{itemize}
+\item \texttt{void UpdateIndex (List<TempIndex> input)}: this method
+  takes input of type List<TempIndex> (Note: Temporary index is a user
+  defined data class, which contains Document ID and List of <keyword
+  ids> as its attributes) and generates an SQL query which updates the
+  INDEX table in the Database.
+
+\item \texttt{List<Integer> Find (List<Integer> QueryKeywordIds)}:
+  this method takes a list of \texttt{Integers} (which corresponds to
+  mapped \texttt{String} query into keyword ids) as input and searches
+  the INDEX table in DB to find all the document ids which are common
+  between for the given query keywords. 
+\end{itemize}
+
+ Figures \ref{fig:indexclassdiagram} and
+ \ref{fig:indexsequencediagram} shows the UML class and sequence
+ diagrams respectively for classes and methods involved in the
+ implementation of the Index data access layer.
+
+\begin{figure}
+  \begin{center}
+	\includegraphics[width=\textwidth,height=!]{indexclassdiagram}
+  \end{center}
+  \caption{Index class diagram}
+  \label{fig:indexclassdiagram}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+	\includegraphics[width=\textwidth,height=!]{indexsequencediagram}
+  \end{center}
+  \caption{Index sequence diagram}
+  \label{fig:indexsequencediagram}
+\end{figure}
+
+% Software Layer
+\subsection{Software Layer}
+
+% Searcher
+\subsubsection{Searcher}
+This class is designed to perform searches for a query entered by the
+user. It contains a method named \texttt{search ()} which is responsible for
+finding all the documents which contained all query keywords.
+
+\begin{itemize}
+\item \texttt{List<Document> search (String)}: given a comma-separated
+  list of search terms, returns the list of documents in which
+  the canonical form of all search terms appear together. This method
+  First calls \texttt{List<int> KeywordDAO.canonize (String)}, to map
+  string query keywords into corresponding list of keyword IDs.  Note:
+  We are assuming that the query will be formed with the set of
+  keywords, which will be a subset of the global keyword set, with which
+  the parsing state machine is constructed.
+
+  Then it calls \texttt{List<Integer> IndexDAO.find (List<querykeywordIDs>)},
+  which will searches in INDEX DB for a list of Common DOC\_IDs
+  corresponding to given query keywords.
+
+  In the end, it calls \texttt{List<Document>
+    DocumentDAO.getDocumentById (DOC\_ID)}. This will map DOC\_ID into
+  a \texttt{Document} from the BLOG table in the database.  
+\end{itemize}
+
+Figures \ref{fig:searcherclassdiagram} and
+\ref{fig:searchersequencediagram} shows the UML class and sequence
+diagrams respectively for classes and methods involved in the
+implementation of the \texttt{Searcher} class.
+
+\begin{figure}
+  \begin{center}
+	\includegraphics[width=\textwidth,height=!]{searcherclassdiagram}
+  \end{center}
+  \caption{Searcher class diagram}
+  \label{fig:searcherclassdiagram}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+	\includegraphics[width=\textwidth,height=!]{searchersequencediagram}
+  \end{center}
+  \caption{Searcher sequence diagram}
+  \label{fig:searchersequencediagram}
+\end{figure}
+
+% State Machine
+\subsubsection{State Machine}
+This class is designed to build the state machine from the list of
+keywords returned by the \texttt{KeywordDAO.getKeywords()} method. This class has a static
+attribute \texttt{initialState} that is the first state of the state machine
+built by the function \texttt{createStateMachine} (initializes with the value
+null). The first state gives us the whole state machine thanks to the
+architecture of a state. Once the state machine is built, the function
+\texttt{getStateMachine} is used to access the initial state stored in the
+class.
+
+Figures \ref{fig:statemachineclassdiagram} and
+\ref{fig:statemachinesequencediagram} shows the UML class and sequence
+diagrams respectively for classes and methods involved in the
+implementation of the \texttt{StateMachineBuilder}.
+
+\begin{figure}
+  \begin{center}
+	\includegraphics[width=\textwidth,height=!]{statemachineclassdiagram}
+  \end{center}
+  \caption{State machine class diagram}
+  \label{fig:statemachineclassdiagram}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+	\includegraphics[width=\textwidth,height=!]{statemachinesequencediagram}
+  \end{center}
+  \caption{State machine sequence diagram}
+  \label{fig:statemachinesequencediagram}
+\end{figure}
+
+% Parser
+\subsubsection{Parser}
+The \texttt{Parser} implements the document parsing algorithm within
+the state machine. By default, it uses the state machine contained
+in the static attribute of \texttt{StateMachineBuilder}. To parse a
+document we need to call the \texttt{parse ()} function inside this
+class, passing it the \texttt{Document} class encapsulating the actual
+text we want to parse. This class returns a \texttt{TemporaryIndex}
+object which encapsulates a reference to the document parsed and the
+keywords found to occur within the document.
+
+Figures \ref{fig:parserclassdiagram} and
+\ref{fig:parsersequencediagram} shows the UML class and sequence
+diagrams respectively for classes and methods involved in the
+implementation of the \texttt{Parser}.
+
+\begin{figure}
+  \begin{center}
+	\includegraphics[width=\textwidth,height=!]{parserclassdiagram}
+  \end{center}
+  \caption{Parser class diagram}
+  \label{fig:parserclassdiagram}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+	\includegraphics[width=\textwidth,height=!]{parsersequencediagram}
+  \end{center}
+  \caption{Parser sequence diagram}
+  \label{fig:parsersequencediagram}
+\end{figure}
 
@@
 processor, 1GB RAM, running a 32 bit Linux 2.6.37 kernel. The Java
-Virtual Machine used was version 1.6.0-24.
-
-\subsection{Aho-Corasick}
-Various tests were conducted to characterize the performance of the
-Aho-Corasick search algorithm:
-
-\begin{itemize}
-\item The time taken to construct the Aho-Corasick state machine was
-  measured for different numbers of keywords. The results of this test
-  is shown in figure \ref{fig:ahostatemachine} and construction time
-  can easily be considered linear with respect to the number of
-  keywords. This is consistent with \cite{RefWorks:103} which proves
-  that the state machine construction algorithm is linearly
-  proportional to the sum of the lengths of the keywords used to
-  construct the state machine.
-
-\item The time taken for the Aho-Corasick state machine to scan
-  different sized corpora was measured. The test was repeated for
-  various state machines, constructed using a 100, 200, and 400
-  keyword set. The results of this test are shown on figure
-  \ref{fig:ahoscan} and indicate that scan time is nearly linear with
-  respect to the size of the corpus and independent of the number of
-  keywords used to generate the state machine. This last result is
-  consistent with \cite{RefWorks:103} which proves that the number of
-  state transitions involved in processing an input string is
-  independent of the number of keywords used to construct the state
-  machine.
-
-\item The time taken to search for different numbers of keywords was
-  measured over different sized corpora. A set of ten keywords was
-  selected randomly and the the corpora searched for ${n \choose k}$
-  enumerated combinations thereof to obtain an average for a
-  particular keyword set size. The results of this test are shown in
-  figure \ref{fig:ahosearch} and indicate that search time is
-  independent of the number of terms used in the search and linear
-  with respect to the size of the corpus. This first result is
-  surprising given that the search process exists as soon as a single
-  instance of each search term is found. One would expect the number
-  of documents scanned completely to increase as the number of search
-  term increases. The small average document size and the fact that
-  the search hit rate is low may account for this phenomenon.
-\end{itemize}
-
-\begin{figure}
-  \begin{center}
-	\includegraphics[width=\textwidth,height=!]{ahostatemachine}
-  \end{center}
-  \caption{Aho-Corasick state machine construction}
-  \label{fig:ahostatemachine}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-	\includegraphics[width=\textwidth,height=!]{ahoscan}
-  \end{center}
-  \caption{Aho-Corasick state machine scanning}
-  \label{fig:ahoscan}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-	\includegraphics[width=\textwidth,height=!]{ahosearch}
-  \end{center}
-  \caption{Aho-Corasick state machine search}
-  \label{fig:ahosearch}
-\end{figure}
-
-\subsection*{Inverted Index}
-Various tests were conducted to characterize the performance of the
-inverted index search algorithm:
-
-\begin{itemize}
-\item The time taken to construct the inverted index was measured for
-  different sized corpora. The results of this test is shown in figure
-  \ref{fig:invscan} and, as expected, indicate a linear relationship
-  between index construction time and the size of the corpora.
-
-\item The time taken to search for different numbers of keywords was
-  measured over different sized corpora. A set of ten keywords was
-  selected randomly and the the corpora searched for ${n \choose k}$
-  enumerated combinations thereof to obtain an average for a
-  particular keyword set size. The results of this test are shown in
-  figure \ref{fig:invsearch}. If the results for one and ten search
-  words are ignored, search times appear to be independent of the
-  number of search terms and close to linear with respect to the size
-  of the corpus. The unusual results for one and ten keywords warrant
-  further investigation.
-\end{itemize}
-
-\begin{figure}
-  \begin{center}
-	\includegraphics[width=\textwidth,height=!]{invscan}
-  \end{center}
-  \caption{Inverted index creation}
-  \label{fig:invscan}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-	\includegraphics[width=\textwidth,height=!]{invsearch}
-  \end{center}
-  \caption{Inverted index search}
-  \label{fig:invsearch}
-\end{figure}
-
-\subsection{Comparison}
-Comparing corpus scan times for Aho-Corasick and inverted index
-implementations, figures \ref{fig:ahoscan} and \ref{fig:invscan}
-respectively, indicates that the inverted index scanner is
-significantly faster than the Aho-Corasick state machine. Given that
-the grammar supported by the JFlex-generated scanner is more expressive
-than that of the Aho-Corasick state machine, one would expect results
-to more comparable. Reviewing the source code of the generated scanner
-indicates a compact and efficient implementation using packed data
-structures to represent the internal state machine. In comparison, the
-Aho-Corasick implementation is relatively unoptimized, closely
-following the pseudo-code in \cite{RefWorks:103}.
-
-Comparing search times for Aho-Corasick and inverted index
-implementations, figures \ref{fig:ahosearch} and \ref{fig:invsearch}
-respectively, indicates that the inverted index implementation has an
-advantage of several orders of magnitude. The initial cost of creating
-the inverted index is significantly larger than the creation of the
-Aho-Corasick state machine however this cost is amortized over each
-subsequent search. The Aho-Corasick implementation rescans the entire
-corpus for each search, as opposed to a scan of the inverted index
-alone. This exercise is a classic demonstration of the impetus behind
-using indexed search techniques for large corpora.
+Virtual Machine used was version 1.6.0-24.....
 
@@
 % Change Bibliography title to 'References'
-\renewcommand\refname{References}
+\renewcommand\bibname{References}
 \bibliography{bibliography}

### 2011-05-15 report/report.tex
Started literature survey

@@
 performance boundaries of a particular system. In addition, as with
-machine learning and classification problems, using a scalar value to
-represent the performance of a system as a whole is misleading and
-does not account for the variety of operating conditions under which
-the system may be expected to perform.
+machine learning and classification problems, using
+a scalar value to represent the performance of a system as a whole is
+misleading and does not account for the variety of operating
+conditions under which the system may be expected to perform.
 
@@
 
-The Aho-Corasick string matching algorithm\cite{RefWorks:103} is a
-kind of dictionary matching search algorithm that constructs a finite
+\textbf{Evalauation}\\
+In the case of machine learning and classification systems, the
+specification of a particular set of misclassification costs or false
+and positive misclassification rates define a particular
+\textit{operating point} of the system. This concept applies equally
+to \textit{information retrieval} systems whereby it is misleading to
+represent the operating characteristics of a particular system by a
+single scalar metric.  Non-scalar evaluation methods try to
+characterize the performance of a system over a range of operating
+conditions. Although this makes direct performance comparisons
+ambiguous, we are more likely to be able to select the algorithm which
+will perform best in the real world by applying domain expertise to
+the results of a non-scalar evaluation \cite{RefWorks:32}.
+
+\textbf{Aho-Corasick}\\
+In order to construct an \textit{inverted index}, documents must be
+scanned to determine \termit{term frequencies}. The Aho-Corasick
+string matching algorithm\cite{RefWorks:103} is a simple and efficient
+text scanning algorithm. The algorithm constructs a finite
 state machine to scan for a given set of keywords. It is, in effect, a
 reduced grammar regular expression parser described in
-\cite{RefWorks:111}.
+\cite{RefWorks:111}. The algorithm is simple and efficient,
+construction time being proportional to the sum of the length of the
+keyword set, and the number of state trasitions required to scan a
+document is independent of the number of the size of the keyword
+set.
+
+\textbf{Parallelization}\\
+Once an objective measure has been established, the performance of an
+information retrieval system can be improved iteratively by making
+various optimizations. These optimizations may take the form of
+algorithmic enhancements, implementation efficiencies, or
+parallelization.
+
+Algorithmic enhancements usually represent the forefront of research
+in the field. Initial work to validate a new concept or algorithm is
+usaully performed in a very controlled environment and on relatively
+small document collections. As the field matures, improvements are
+harder fought and often involve increasingly complex algorithms
+and corpus approximation methods into which it becomes more difficult
+to \textit{fold-in} new documents. In addressing the considerable
+computation cost of re-calculating \textit{LSI} approximations,
+\cite{RefWorks:101} introduced new update methods which saw
+improved average precisions but were forced to conclude that the
+improvements were at a computational premium over previous methods.
+
+Enhancement of implementation efficiency in
+\cite{RefWorks:96} saw a six-fold timing improvment in \textit{LSI}
+searches on large document collections. The same study achieved nearly
+180 times improvement through the use of parallelization. The study
+made no algorithmic improvements and serves to re-inforce the fact that
+parallelization is an important mechanism employed by large commercial
+systems to cope with the ever increasing bodies of data that must be
+processed.
 

### 2011-05-16 report/report.tex
First draft of literature survey complete

--- 2011-05-13 report/report.tex
@@
 quality of a particular algorithm or system and compare it to
-others. So often, information retrieval algorithms lack an analytical
+others. Often, information retrieval algorithms lack an analytical
 model that would allow one to make formal statements as to the
-performance boundaries of a particular system. In addition, as with
-machine learning and classification problems, using a scalar value to
-represent the performance of a system as a whole is misleading and
-does not account for the variety of operating conditions under which
-the system may be expected to perform.
+performance characteristics of a particular system. In addition, as with
+machine learning and classification problems, using
+a scalar value to represent the performance of a system as a whole is
+misleading and does not account for the variety of operating
+conditions under which the system may be expected to perform.
 
@@
       storage requirements evolve with the size of the collection
-      and/or degree of concurrency employed ?
+      and/or degree of concurrency ?
 \end{itemize}
@@
 To drive information retrieval research in one direction or another,
-it would be useful to have at ones disposal a physical implementation
+it would be useful to have a physical implementation
 framework within which one could empirically explore the operating
-parameters of a particular algorithm and/or technique along the lines
+characteristics of a particular algorithm and/or technique along the lines
 described above. Such a system would allow researchers to investigate
 how the system responds to variation in any one of these operating
-parameters. Does the query speed increase as more computational
-capacity is added ? How do storage requirements vary with the size of
-the document collection ? The aim of this project is to create such an
+parameters. The aim of this project is to create such an
 information retrieval \textit{test-bed}.
@@
 architected such that the specific algorithm could easily be changed
-without major rework. This paper presents a survey of existing
+without significant rework. This paper presents a survey of existing
 literature in support of our test-bed concept, as well as the specific
-algorithmic techniques embodied in the reference implementation. In
+algorithms employed in the reference implementation. In
 addition, a detailed architectural overview of the system is
@@
 
-The Aho-Corasick string matching algorithm\cite{RefWorks:103} is a
-kind of dictionary matching search algorithm that constructs a finite
+\subsection{Indexing}
+\textit{Information Retrieval} is a fast becoming the dominant form of
+information access, overtaking traditional relational database style
+searching \cite{RefWorks:109}. The basis of most document retrieval
+systems is the \textit{term-document} incidence matrix. This matrix is
+typically sparse and more efficiently represented as an
+\textit{inverted-index} which maps terms to the parts of a document
+where they occur.
+
+In order to construct an \textit{inverted index}, documents must be
+scanned to determine \termit{term frequencies}. The Aho-Corasick
+string matching algorithm\cite{RefWorks:103} is a simple and efficient
+text scanning algorithm. The algorithm constructs a finite
 state machine to scan for a given set of keywords. It is, in effect, a
-reduced grammar regular expression parser described in
-\cite{RefWorks:111}.
+reduced grammar regular expression parser of the type described in
+\cite{RefWorks:111}. The algorithm is simple and efficient,
+construction time being proportional to the sum of the length of the
+keyword set, and the number of state transitions required to scan a
+document is independent of the number of the size of the keyword
+set.
+
+\subsection{Evaluation}
+In the case of machine learning and classification systems, the
+specification of a particular set of misclassification costs or false
+and positive misclassification rates define a particular
+\textit{operating point} of the system.
+
+This concept applies equally to \textit{information retrieval} systems
+whereby it is misleading to represent the operating characteristics of
+a particular system with a scalar metric. In addition, the evaluation
+of information retrieval systems suffers from the the inherent
+subjectivity with respect to \textit{relevance}
+\cite{RefWorks:72}. Non-scalar evaluation methods try to characterize
+the performance of a system over a range of operating
+conditions. Although this makes direct performance comparisons
+ambiguous, we are more likely to be able to select the algorithm which
+will perform best in the real world by applying domain expertise to
+the results of a non-scalar evaluation \cite{RefWorks:32}.
+
+\subsection{Parallelization}
+Once an objective measure has been established, the performance of an
+information retrieval system can be improved iteratively by making
+various optimizations. These optimizations may take the form of
+algorithm enhancements, implementation efficiencies, or
+parallelization.
+
+Algorithm enhancements usually represent the forefront of research in
+the field. Work to validate a new concept or algorithm is usually
+performed in a very controlled environment and on relatively small
+document collections. As the field matures, improvements are harder
+fought and often involve increasingly complex algorithms and corpus
+approximation methods into which it becomes more difficult to
+\textit{fold-in} new documents. In addressing the considerable
+computation cost of re-calculating \textit{LSI} \cite{RefWorks:102}
+approximations, \cite{RefWorks:101} introduced new update methods
+which saw improved average precisions but were forced to conclude that
+the improvements were at a computational premium over previous
+methods. A semi-discrete matrix decomposition for LSI, \textit{SDD},
+performs as well as the original SVD-based LSI method for equal query
+times and uses less storage \cite{RefWorks:99}. The decomposition time
+of the mew method, however, is significantly larger than the original.
+
+Enhancement of implementation efficiency in \cite{RefWorks:96} saw a
+six-fold timing improvement in \textit{LSI} searches on large document
+collections. The same study achieved nearly 180 times improvement
+through the use of parallelization. The study made no algorithmic
+improvements and the result is consistent with increased use of
+parallelization by large commercial systems to cope with the ever
+increasing bodies of data that must be processed.
+
+\textit{Map-Reduce}, popularized by Google, is emerging as an
+important programming model for large-scale data-parallel applications
+such as web indexing \cite{RefWorks:113}. Fundamentally, Map-Reduce
+breaks up a computation into smaller tasks which are made to run in
+parallel on different physical machines. The \textit{map} and
+\textit{reduce} phases are inspired by the functions of the same name
+used in \textit{functional programming}. The model scales well to large
+clusters mostly because of the decoupled nature of the tasks
+themselves. An initial \textit{map} phase breaks the data into
+  separate chunks and assigns them to a task. The \textit{reduce}
+  phase assembles the results of the map tasks to produce the final
+  result.
+
+\textit{Hadoop} is an open source implementation of the MapReduce
+model, developed mostly by Yahoo, and now maintained by the Apache
+Software Foundation. Hadoop can be configured to automatically create
+speculative copies of slow running map tasks
+(stragglers). This facility is intended as a pre-emptive failure
+recovery mechanism as well as a means of optimizing overall
+throughput. The metrics used to identify stragglers as well as the
+algorithm for determining the target node on which speculative copies
+of a task should be launched, is examined in \cite{RefWorks:128}. The
+paper introduces an alternative scheduling algorithm, \textit{LATE},
+to address excessive task speculation issues with the native Hadoop
+scheduler in a heterogeneous hardware environment. Experiments
+comparing the native Hadoop scheduler, a non-speculative scheduling
+strategy, and LATE were conducted using hosted virtual machines (VMs)
+in the Amazon EC2 environment as well as a dedicated local
+cluster. Variance in node performance was introduced by varying the
+distribution of VMs across physical machines and also introducing
+network and disk transfer loads. Results show speculative execution
+(native Hadoop and LATE) outperforming non-speculative execution and
+LATE outperforming native Hadoop in most cases.
 

### 2011-06-06 report/report.tex
First draft of literature survey complete
Started final merge

@@
 \usepackage{graphicx}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 \usepackage{setspace}                   
@@
 \title{CS6999 Web Data and Text Mining:\\Parallel Text Indexer Project Report}
-\author{Damien DuBios\\Justin Kamerman 3335272\\Raman Singh}
+\author{Damien DuBois\\Justin Kamerman 3335272\\Ramanpreet Singh 3395913}
 \date{\today}
@@
 times and uses less storage \cite{RefWorks:99}. The decomposition time
-of the mew method, however, is significantly larger than the original.
+of the new method, however, is significantly larger than the original.
 
@@
 %----------------------------------------
-% Motivation
-%----------------------------------------
-\section{Motivation}
-\label{sec:motivation}
-
-%----------------------------------------
 % Research Plan
@@
 
+The research design and methods section should include the following:
+
+    an overview of the experimental design
+
+    a detailed description of specific methods to be employed to accomplish the specific aims
+
+    a detailed discussion of the way in which the results will be collected, analyzed, and interpreted
+
+    a projected sequence or timetable (work plan)
+
+    a description of any new methodology used and why it represents an improvement over the existing ones
+
+    a discussion of potential difficulties and limitations and how these will be overcome or mitigated
+
+    expected results, and alternative approaches that will be used if unexpected results are found
+
+    precautions to be exercised with respect to any procedures, situations, or materials that may be hazardous to personnel or human subjects
+
+Suggestion
+
+    Number the sections in this part of the application to correspond to the numbers of the Specific Aims.
+
+    Give sufficient detail. Do not assume that the reviewers will know how you intend to proceed.
+
+    Avoid excessive experimental detail by referring to publications that describe the methods to be employed.
+
+    Publications cited should be by the applicants, if at all possible.
+Citing someone else's publication establishes that you know what method to use, but citing your own (or that of a collaborator) establishes that the applicant personnel are experienced with the necessary techniques.
+
+    If relevant, explain why one approach or method will be used in preference to others. This establishes that the alternatives were not simply overlooked. Give not only the "how" but the "why."
+
+    If employing a complex technology for the fast time, take extra care to demonstrate familiarity with the experimental details and potential pitfalls.
+Add a coinvestigator or consultant experienced with the technology, if necessary.
+
+    Document proposed collaborations and offers of materials or reagents of restricted availability with letters from the individuals involved.
+
+\bigskip
+The whole idea of the project belongs in being able to reduce the time
+to build the index. As our system is building the inverted index by
+parsing all the post into the state machine, we can expect that the
+complexity will be proportional to the whole number of characters of
+all the post we want to build the index for. As a result changing the
+number of Posts, the number of characters inside of each post should
+have a big impact in our time complexity. Besides, we should also have
+a time needed at the very beginning of the running to build the state
+machine. We can expect that the building of the state machine needs a
+time proportional to the sum of the size of the keywords.
+
+The system proposed in this paper doesn’t take care of various
+compression techniques for storing the index. To do that, many
+techniques have been explored in the literature. We will just
+highlight the different sizes of our indexes according to different
+parameters.
+
+Finally we will do a time comparison between our way of building the
+inverted index and the naive way.  All the time complexities tests are
+done on a single computer and it is important to understand that all
+these processes can be parallelized. A future work would be to install
+the system on multiple computers and explore the gain in time by
+running the code simultaneously on these computers.
+
 %----------------------------------------
-% Design Component
+% Design
 %----------------------------------------
-\section{Design Component}
-\label{sec:designcomponent}
+\section{Design}
+\label{sec:design}
 A black-box functional view of the indexing system is show in figure
@@
 The following subsections detail the implementation of the
-architectural layers identified in section \ref{sec:designcomponent}.
+architectural layers identified in section \ref{sec:design}.
 
@@
 %----------------------------------------
+% Number of Keywords
+%----------------------------------------
+\subsection{Number of Keywords}
+To be able to study the impact of different number of keywords, we
+built a keywords database with 10.000 most popular English
+words. These 10.000 words come from:
+http://www.mit.edu/~ecprice/wordlist.10000 All the experiments in this
+part are done with 50.000 posts randomly downloaded from internet. The
+idea is to get some heterogeneous length and type of post such that it
+is realistic.  The experiment shown here consist on building the index
+for different number of keywords from the 10.000 keywords database and
+explore the time and space complexity for the building of the state
+machine and the building of the index once the state machine is built.
+
+\subsubsection{Time Complexity}
+\begin{itemize}
+\item \textbf{State Machine Construction:}
+  The time taken to construct the Aho-Corasick state machine was
+  measured for different numbers of keywords. The results of this test
+  is shown in figure \ref{fig:numkeywordstimecomplexbuildsm} and construction time
+  can easily be considered linear with respect to the number of
+  keywords. This is consistent with \cite{RefWorks:103} which proves
+  that the state machine construction algorithm is linearly
+  proportional to the sum of the lengths of the keywords used to
+  construct the state machine.
+
+  \begin{figure}
+    \begin{center}
+      \includegraphics[width=\textwidth,height=!]{numkeywordstimecomplexbuildsm}
+    \end{center}
+    \caption{Time to build Aho-Corasick state machine as a function of
+      the number of keywords. Time taken is shown in miliseconds on the
+      y-axis and the number of keywords on the x-axis}
+    \label{fig:numkeywordstimecomplexbuildsm}
+  \end{figure}
+
+\item \textbf{Index Construction:} 
+  The time taken for a state machines to build an inverted index of our
+  corpus was measured. The test was repeated using state machines
+  constructed with a varying number of keywords and the results plotted
+  in figure \ref{fig:numkeywordstimecomplexbuildindex}. According to
+  \cite{RefWorks:103} which proves that the number of state transitions
+  involved in processing an input string is independent of the number of
+  keywords used to construct the state machine, we would not expect the
+  time complexity to increase with the size of the keyword set. However,
+  as can be seen in \ref{fig:numkeywordstimecomplexbuildindex}, the time
+  to build the index increases logarithmically with the number of
+  keywords. This is likely due to the fac that as the number of keywords
+  increases, so does the size of the index and the number of database
+  interacts required to store and update the index.
+
+  \begin{figure}
+    \begin{center}
+      \includegraphics[width=\textwidth,height=!]{numkeywordstimecomplexbuildindex}
+    \end{center}
+    \caption{Time to build the inverted index as a function of the
+      number of keywords. Time taken is shown in miliseconds on the
+      y-axis and the number of keywords on the x-axis}
+    \label{fig:numkeywordstimecomplexbuildindex}
+  \end{figure}
+
+\end{itemize}
+
+\subsubsection{Space Complexity}
+As we would have expected it, the space used by the index in the
+database is proportional to the number of words we are using. It also
+means that our words have the same probability at the beginning or at
+the end of our 12,000 words dictionnary. This result is reflected in
+figure \ref{fig:numkeywordsspacecomplexbuildindex}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth,height=!]{numkeywordsspacecomplexbuildindex}
+  \end{center}
+    \caption{Size of index as a function of the number of
+      keywords. The number of database rows used to store the index is
+      shown on the y-axis and the number of keywords on the x-axis} 
+    \label{fig:numkeywordsspacecomplexbuildindex}
+\end{figure}
+
+%----------------------------------------
+% Size of Corpus
+%----------------------------------------
+\subsection{Size of Corpus}
+To be able to study the impact of different number of Posts, we built
+a posts database with 400.000 posts randomly downloaded from
+internet. Then the keywords used to build the index are simply the
+first 5.000 from the previous keywords database.
+
+The experiment shown here consist on building the index for different
+number of Posts and explore the time and space complexity for the
+building of the state machine and the building of the index once the
+state machine is built. We can expect that the time to build the State
+machine won’t change with the number of posts and the time to build
+the index will increase almost linearly with the number of posts.
+
+\subsubsection{Time Complexity}
+As we were expecting, the time to build the index with the state
+machine strategy is linear with respect to the number of post, as
+depicted in figure \ref{fig:corpsizetimecomplexbuildindex}. However, to be able
+to perform this experiment we had to run the indexer program with
+batches of  10,000 posts due to Java heap space problems. As a result, the state
+machine is rebuilt every 10,000 posts.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth,height=!]{corpsizetimecomplexbuildindex}
+  \end{center}
+    \caption{Time to build the inverted index as a function of the
+      size of the corpus. Time taken is shown in miliseconds on the
+      y-axis and the number of documents in the collection on the
+      x-axis} 
+    \label{fig:corpsizetimecomplexbuildindex}
+\end{figure}
+
+\subsubsection{Space Complexity}
+As expected, the space used to sptore the index in the database is
+proportional to the size of the corpus. This result is shown in figure
+\ref{fig:corpsizespacecomplexbuildindex}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth,height=!]{corpsizespacecomplexbuildindex}
+  \end{center}
+    \caption{Size of index as a function of the size of the
+      corpus. The number of database rows used to store the index is
+      shown on the y-axis and the number of documents in the
+      collection on the x-axis} 
+    \label{fig:corpsizespacecomplexbuildindex}
+\end{figure}
+
+%----------------------------------------
+% Document Size
+%----------------------------------------
+\subsection{Document Size} 
+As we have seen before, the time to build the index is almost
+proportional with the number of post. Theoretically this time should
+be proportional to the total length of all posts. Let’s focus on this
+aspect and study the impact of the length of the post on the time to
+build the state machine. To do that we built a 5.000 posts database
+with posts that are all composed of 20.000 characters. Then, we ran
+our system on different length of substring of these posts.
+
+\subsubsection{Time Complexity}
+\begin{itemize}
+\item \textbf{State Machine Construction:} 
+  As expected, state machine construction time
+  does not depend on the length of the documents. Nonetheless it is
+  interesting to compare figures \ref{fig:docsizetimecomplexbuildsm}
+  and \ref{fig:docsizetimecomplexbuildindex}, showing the construction
+  times for the state machine and inverted index relative to the
+  average document size. The comparison shows that state machine
+  construction time is several orders of magnitude smaller than index
+  construction and negligible by comnparison. Our state machine is
+  built in 770 ms on average for 1000 keywords with a standard
+  deviation of 18ms.
+
+  \begin{figure}
+    \begin{center}
+      \includegraphics[width=\textwidth,height=!]{docsizetimecomplexbuildsm}
+    \end{center}
+    \caption{Time to build Aho-Corasick state machine as a function of
+      the average document size. Time taken is shown in miliseconds on the
+      y-axis and the average document size in characters on the x-axis}
+    \label{fig:docsizetimecomplexbuildsm}
+  \end{figure}
+
+\item \textbf{Index Construction:} 
+  As expected the time to build the indexer is almost linear with
+  respect to the average document size. This relationship is shown in
+  figure \ref{fig:docsizetimecomplexbuildindex}.
+
+  \begin{figure}
+    \begin{center}
+      \includegraphics[width=\textwidth,height=!]{docsizetimecomplexbuildindex}
+    \end{center}
+    \caption{Time to build an inverted index as a function of the
+      average document size. Time taken is shown in miliseconds on the
+      y-axis and the average document size in characters on the x-axis} 
+    \label{fig:docsizetimecomplexbuildindex}
+  \end{figure}
+
+\end{itemize}
+
+\subsubsection{Space Complexity}
+Here we are studying the number of records in our index database as a
+function of the average document size. As expected, the larger the documents
+the more records created to represent the index. However, this is not linear
+since if a word occurs many times in the same post we just save it
+once in the database. This relationship is shown in figure
+\ref{fig:docsizespacecomplexbuildindex}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth,height=!]{docsizespacecomplexbuildindex}
+  \end{center}
+  \caption{Size of index as a function of the average document
+    size. The number of database rows used to store the index is
+      shown on the y-axis and the average document size in characters
+      on the x-axis}  
+  \label{fig:docsizespacecomplexbuildindex}
+\end{figure}
+
+%----------------------------------------
+% Comparison with Naive Index Builder
+%----------------------------------------
+\subsection{Comparison with Na\"{a}ve Index Builder}
+This part focuses on the improvement that our system with respect to
+time complexity in comparison with a na\"{i}e index builder. The na\"{i}ve
+indexer builder has the same input as our system i.e. a list of
+keywords and a list of documents, and the same output i.e. an inverted
+index corresponding to the keywords and documents. Algorithm
+\ref{alg:invertedindex} is used in principal for both implementations.
+
+\begin{algorithm}
+\caption{Build inverted index}
+\label{alg:invertedindex}
+\begin{algorithmic}
+  \FORALL{$keyword$}
+  \STATE Build a vector $v$
+  \FORALL {$document$}
+  \IF {$keyword$ \epsilon \; $document$}
+  \STATE Add the row (keyword.id, document.id) into $v$
+  \ENDFOR
+  \STATE Save $v$ into the database
+  \ENDFOR
+\end{algorithmic}
+\end{algorithm}
+
+As we can see this algorithm has a complexity that directly depends on
+the number of keywords n, the number of document m and the length of
+document since we have to check in each document for each
+keyword. Moreover, we decided to limit as much as possible the
+accesses to the database since it is really time consuming in
+Java. That’s why we do one access to the database for each keywords
+and not for each document.
+
+Note: In the state machine we do as much accesses to the database as
+the number of document (in the worst case meaning we find at least one
+keyword per document).
+
+Let’s compare the results for this algorithm with our system results
+for different number of keywords and posts.
+
+\subsubsection{Analytical}
+
+\subsubsection{Experimental}
+
+%----------------------------------------
 % Conclusions
@@
 
+The experimental results tend to prove what we could expect from
+theory. Our system is faster than the naive indexer whatever
+parameters we are changing. In particular when the number of document
+is becoming really important, the State machine approach prove to be
+very efficient. Moreover, as soon as we are dealing with a number of
+keywords that can be representative of industrial needs, the state
+machine prove also to be far faster than the naive indexer.
+
+Future work on this strategy would be to compare this approach with
+less naive strategies to build an index. Moreover, as said in the
+introduction, it would probably be interesting to add compression
+methods to our algorithm in order to reduce the index size in the
+database.
+
 %----------------------------------------
@@
 %----------------------------------------
-% Change Bibliography title to 'References'
+% changes default name Bibliography to References
 \renewcommand\bibname{References}

### 2011-06-06 report/report.tex
Finished merging results except for graphs

@@
 
-    an overview of the experimental design
-
-    a detailed description of specific methods to be employed to accomplish the specific aims
-
-    a detailed discussion of the way in which the results will be collected, analyzed, and interpreted
-
-    a projected sequence or timetable (work plan)
-
-    a description of any new methodology used and why it represents an improvement over the existing ones
-
-    a discussion of potential difficulties and limitations and how these will be overcome or mitigated
-
-    expected results, and alternative approaches that will be used if unexpected results are found
-
-    precautions to be exercised with respect to any procedures, situations, or materials that may be hazardous to personnel or human subjects
-
-Suggestion
+\begin{itemize}
+    \item an overview of the experimental design
+
+    \item a detailed description of specific methods to be employed to accomplish the specific aims
+
+    \item a detailed discussion of the way in which the results will be collected, analyzed, and interpreted
+
+    \item a projected sequence or timetable (work plan)
+
+    \item a description of any new methodology used and why it represents an improvement over the existing ones
+
+    \item a discussion of potential difficulties and limitations and how these will be overcome or mitigated
+
+    \item expected results, and alternative approaches that will be used if unexpected results are found
+
+    \item precautions to be exercised with respect to any procedures, situations, or materials that may be hazardous to personnel or human subjects
+\end{itemize}
+
+Suggestion:
 
@@
     Document proposed collaborations and offers of materials or reagents of restricted availability with letters from the individuals involved.
-
-\bigskip
-The whole idea of the project belongs in being able to reduce the time
-to build the index. As our system is building the inverted index by
-parsing all the post into the state machine, we can expect that the
-complexity will be proportional to the whole number of characters of
-all the post we want to build the index for. As a result changing the
-number of Posts, the number of characters inside of each post should
-have a big impact in our time complexity. Besides, we should also have
-a time needed at the very beginning of the running to build the state
-machine. We can expect that the building of the state machine needs a
-time proportional to the sum of the size of the keywords.
-
-The system proposed in this paper doesn’t take care of various
-compression techniques for storing the index. To do that, many
-techniques have been explored in the literature. We will just
-highlight the different sizes of our indexes according to different
-parameters.
-
-Finally we will do a time comparison between our way of building the
-inverted index and the naive way.  All the time complexities tests are
-done on a single computer and it is important to understand that all
-these processes can be parallelized. A future work would be to install
-the system on multiple computers and explore the gain in time by
-running the code simultaneously on these computers.
 
@@
 \label{sec:experimentresults}
-FIX: All tests were run on a Dell laptop with an Intel Core Duo 2GHz
+The whole idea of the project belongs in being able to reduce the time
+to build the index. As our system is building the inverted index by
+parsing all the post into the state machine, we can expect that the
+complexity will be proportional to the whole number of characters of
+all the post we want to build the index for. As a result changing the
+number of Posts, the number of characters inside of each post should
+have a big impact in our time complexity. Besides, we should also have
+a time needed at the very beginning of the running to build the state
+machine. We can expect that the building of the state machine needs a
+time proportional to the sum of the size of the keywords.
+
+The system proposed in this paper doesn’t take care of various
+compression techniques for storing the index. To do that, many
+techniques have been explored in the literature. We will just
+highlight the different sizes of our indexes according to different
+parameters.
+
+Finally we will do a time comparison between our way of building the
+inverted index and the naive way.  All the time complexities tests are
+done on a single computer and it is important to understand that all
+these processes can be parallelized. A future work would be to install
+the system on multiple computers and explore the gain in time by
+running the code simultaneously on these computers.
+
+All tests were run on a single Intel Core 2 Duo 2GHz
 processor, 1GB RAM, running a 32 bit Linux 2.6.37 kernel. The Java
-Virtual Machine used was version 1.6.0-24.....
+Virtual Machine used was version 1.6.0-24. The database used was MySQL
+x.x, running on ....
 
@@
 %----------------------------------------
-\subsection{Comparison with Na\"{a}ve Index Builder}
+\subsection{Comparison with Na\"{i}ve Index Builder}
 This part focuses on the improvement that our system with respect to
@@
 As we can see this algorithm has a complexity that directly depends on
-the number of keywords n, the number of document m and the length of
-document since we have to check in each document for each
-keyword. Moreover, we decided to limit as much as possible the
+the number of keywords $n$, the number of document $m$, and the length
+of document since we have to check in each document for each
+keyword. Moreover, we decided to limit, as much as possible, the
 accesses to the database since it is really time consuming in
 Java. That’s why we do one access to the database for each keywords
-and not for each document.
-
-Note: In the state machine we do as much accesses to the database as
-the number of document (in the worst case meaning we find at least one
-keyword per document).
-
-Let’s compare the results for this algorithm with our system results
-for different number of keywords and posts.
-
-\subsubsection{Analytical}
-
-\subsubsection{Experimental}
+and not for each document. In the state machine we do as much accesses
+to the database as the number of document (in the worst case meaning
+we find at least one keyword per document). In the following
+subsections we compare the results for this algorithm with our system
+results for different number of keywords and posts.
+
+\subsubsection{Complexity Analysis}
+For the na\"{i}ve algorithm described above, the theoretical worst
+case time complexity relationship is shown in equation
+\ref{eqn:timecomplexworstcase}.
+
+\begin{equation}
+\label{eqn:timecomplexworstcase}
+Time \propto keyNum * ( docNum * maxLength + conn)
+\end{equation}
+
+where \(keyNum\) is the number of keywords, \(docNum\) is the number
+of documents, \(maxLength\) is the maximum document length, and
+\(conn\) is the time taken to create a connection to the database. For
+the state machine algorithm the theoretical complexity relationship is
+given by 
+equation \ref{eqn:timecomplexsm}.
+
+\begin{equation}
+\label{eqn:timecomplexsm}
+Time \propto docNum * ( maxLength + conn)
+\end{equation}
+
+\subsubsection{Document Size}
+For this experiment we fixed the number of keywords to 1,000 and the
+number of documents to 5,000 and we ran our code on a database
+composed of document of different length. The experiment results are
+shown in figure \ref{fig:naivedocumentsize}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth,height=!]{naivedocumentsize}
+  \end{center}
+  \caption{Time to build an inverted index as a function of the
+      average document size. Time taken is shown in miliseconds on the
+      y-axis and the average document size in characters on the x-axis}
+  \label{fig:naivedocumentsize}
+\end{figure}
+
+As expected, the curves for both our indexer and the naive indexer are
+almost linear since for the naive implementation:
+
+\[Time \propto (keyNum * docNum) * maxLength + conn * keyNum\]
+
+and for our Aho-Corasick state machine method:
+
+\[Time \propto docNum * maxLength + conn * docNum\]
+
+We can observe that for a small length of document naive method is
+faster since \(keyNum\) is smaller than \(docNum\). For \(docLength = 1300\)
+our indexer is far more efficient. To see this effect more clearly,
+experiment results are plotted on a logarithmic scale in figure
+\ref{fig:naivedocumentsizelog}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth,height=!]{naivedocumentsizelog}
+  \end{center}
+  \caption{Time to build an inverted index as a function of the
+      average document size. Time taken is shown in miliseconds on the
+      y-axis and the average document size in characters on the
+      logarithmic x-axis}
+  \label{fig:naivedocumentsizelog}
+\end{figure}
+
+\subsubsection{Number of Keywords}
+As expected, the two implementations exhibit a linear response to the
+number of keywords used. In the na\"{i}ve implmementation:
+
+\[Time \propto (keyNum * maxLength) * docNum + conn * keyNum\]
+
+and in our Aho-Corasick state machine method:
+
+\[Time \propto (maxLength + conn) * docNum \]
+
+However, we would have expected the state machine method not to depend
+on the number of keyword. The reason why it’s not the case is because
+when we don’t have any keywords in a document, we don’t access the
+database. Knowing that the database access is very time consuming,
+increasing the number of keyword increases the probability of finding a
+keyword into each document and so increase the time complexity.
+
+The experiment results are shown in figure \ref{fig:naivedocumentsize}
+and on a logarithmic scale in figure \ref{fig:naivedocumentsizelog}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth,height=!]{naivenumkeywords}
+  \end{center}
+  \caption{Time to build an inverted index as a function of the
+      number of keywords. Time taken is shown in miliseconds on the
+      y-axis and the number of keywords on the x-axis}
+  \label{fig:naivenumkeywords}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth,height=!]{naivenumkeywordslog}
+  \end{center}
+  \caption{Time to build an inverted index as a function of the
+      number of keywords. Time taken is shown in miliseconds on the
+      y-axis and the number of keywords on the logarithmic x-axis}
+  \label{fig:naivenumkeywordslog}
+\end{figure}
+
+\subsubsection{Size of Corpus}
+As expected, the time complexity is linearly related to the number of
+documents being indexed. The experiment results are shown in figure
+\ref{fig:naivesizecorpus} and on a logarithmic scale in figure
+\ref{fig:naivesizecorpuslog}. For the na\"{i}ve implementation:
+
+\[Time \propto (docNum * maxLenght + conn) * keyNum\]
+
+and in our Aho-Corasick state machine method:
+
+\[Time \propto docNum * maxLenght + conn * docNum\]
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth,height=!]{naivesizecorpus}
+  \end{center}
+  \caption{Time to build an inverted index as a function of the
+      number of documents being indexed. Time taken is shown in
+      miliseconds on the y-axis and the number of documents on the
+      x-axis} 
+  \label{fig:naivesizecorpus}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth,height=!]{naivesizecorpuslog}
+  \end{center}
+  \caption{Time to build an inverted index as a function of the
+      number of documents being indexed. Time taken is shown in
+      miliseconds on the y-axis and the number of documents on the
+      logarithmic x-axis} 
+  \label{fig:naivesizecorpuslog}
+\end{figure}
 
@@
 \label{sec:conclusions}
-
 The experimental results tend to prove what we could expect from
-theory. Our system is faster than the naive indexer whatever
-parameters we are changing. In particular when the number of document
-is becoming really important, the State machine approach prove to be
-very efficient. Moreover, as soon as we are dealing with a number of
-keywords that can be representative of industrial needs, the state
-machine prove also to be far faster than the naive indexer.
-
-Future work on this strategy would be to compare this approach with
-less naive strategies to build an index. Moreover, as said in the
-introduction, it would probably be interesting to add compression
-methods to our algorithm in order to reduce the index size in the
-database.
+theory. Our system is faster than the na\"{i}ve indexer with respect
+to the number of keywords being indexed, size of the document
+collection, and average document size. In particular, our indexer
+implementation performs relatively well with respect to the number of
+documents being indexed. Moreover, as soon as we are dealing with a number of
+keywords that can be representative of real-world needs, the state
+machine implementation proves also to be significantly faster than the
+naive indexer.
+
+One of the primary goals of this research project was to leverage the
+effect of parallelization in reducing the time to index a document
+collection. Due to time constraints, this was not done but the
+resulting implementation lays a strong foundation for further research
+in this area. Other avenues of interest include, but are not limited
+to:
+
+\begin{itemize}
+  \item Compare the Aho-Corasick state machine indexer implementation
+    to less na\"{i}ve indexer implementations.
+
+    \item Use compression techniques to our algorithm in order to
+      reduce the index size in the database.
+
+    \item Study the time performance improvements possible deploying
+      the Aho-Corasick state machine indexer within a commercial
+      \textit{MapReduce} framework such as \textit{Hadoop}.
+\end{itemize}
 

### 2011-06-07 report/report.tex
First draft

@@
 \textit{fold-in} new documents. In addressing the considerable
-computation cost of re-calculating \textit{LSI} \cite{RefWorks:102}
-approximations, \cite{RefWorks:101} introduced new update methods
-which saw improved average precisions but were forced to conclude that
-the improvements were at a computational premium over previous
-methods. A semi-discrete matrix decomposition for LSI, \textit{SDD},
-performs as well as the original SVD-based LSI method for equal query
-times and uses less storage \cite{RefWorks:99}. The decomposition time
-of the new method, however, is significantly larger than the original.
+computation cost of re-calculating \textit{Latent Semantic Indexing}
+(LSI) \cite{RefWorks:102} approximations, \cite{RefWorks:101}
+introduced new update methods which saw improved average precisions
+but were forced to conclude that the improvements were at a
+computational premium over previous methods. A semi-discrete matrix
+decomposition for LSI, \textit{SDD}, performs as well as the original
+SVD-based LSI method for equal query times and uses less storage
+\cite{RefWorks:99}. The decomposition time of the new method, however,
+is significantly larger than the original.
 
@@
 \label{sec:researchplan}
-
-The research design and methods section should include the following:
+The primary goals of this research project is to construct an
+information retrieval framework within which specific implementations
+can be evaluated under various operating conditions. To initiate the
+synthesis of this general framework, a reference implementation will
+be built, based on the notion of a simple inverted index.
+
+The software architecture of the reference implementation will be
+carefully documented and follow established design patterns in order
+to facilitate the future construction of alternative document search
+schemes within the same framework. The framework will be designed to
+be executed in a distributed deployment such that it may be used to
+evaluate the effect of parallelization on a particular algorithm.
+
+The performance of the reference implementation will be tested with
+respect to the following operating parameters and compared to analytic
+expectations:
 
 \begin{itemize}
-    \item an overview of the experimental design
-
-    \item a detailed description of specific methods to be employed to accomplish the specific aims
-
-    \item a detailed discussion of the way in which the results will be collected, analyzed, and interpreted
-
-    \item a projected sequence or timetable (work plan)
-
-    \item a description of any new methodology used and why it represents an improvement over the existing ones
-
-    \item a discussion of potential difficulties and limitations and how these will be overcome or mitigated
-
-    \item expected results, and alternative approaches that will be used if unexpected results are found
-
-    \item precautions to be exercised with respect to any procedures, situations, or materials that may be hazardous to personnel or human subjects
+\item Number of index keywords
+\item Number of documents in collection
+\item Average size of documents
 \end{itemize}
 
-Suggestion:
-
-    Number the sections in this part of the application to correspond to the numbers of the Specific Aims.
-
-    Give sufficient detail. Do not assume that the reviewers will know how you intend to proceed.
-
-    Avoid excessive experimental detail by referring to publications that describe the methods to be employed.
-
-    Publications cited should be by the applicants, if at all possible.
-Citing someone else's publication establishes that you know what method to use, but citing your own (or that of a collaborator) establishes that the applicant personnel are experienced with the necessary techniques.
-
-    If relevant, explain why one approach or method will be used in preference to others. This establishes that the alternatives were not simply overlooked. Give not only the "how" but the "why."
-
-    If employing a complex technology for the fast time, take extra care to demonstrate familiarity with the experimental details and potential pitfalls.
-Add a coinvestigator or consultant experienced with the technology, if necessary.
-
-    Document proposed collaborations and offers of materials or reagents of restricted availability with letters from the individuals involved.
+In addition, the reference implementation will be compared against a
+na\"{i}ve implementation of the inverted indexer. The na\"{i}ve
+implementation provided a baseline for further evaluation.
 
@@
   \begin{center}
-        \includegraphics[width=\textwidth,height=!]{overallclassdiagram}
+        \includegraphics[angle=90,width=\textwidth,height=!]{overallclassdiagram}
   \end{center}
@@
 index). The term occurrences for each document are used to augment an
-existing index for documents already processed. The document
+existing index of documents already processed. The document
 collection, index, and keyword list are maintained in a single MySQL
@@
 potential to become a throttle point as the system is scaled to larger
-and faster-evolving document collections, and is required to service a
-higher rate of queries. In order to scale the current implementation
-to large document collections, a more efficient mechanism would be
-required in this respect.
+and more rapidly evolving document collections, and is required to
+service a higher rate of queries. In order to scale the current
+implementation to large document collections, a more efficient
+mechanism would be required in this respect.
 
@@
   processing is not duplicated on other nodes. In this scheme, a
-  document is considered processed to all other processes once
-  retrieved from the database. This may cause problems from a fault
-  recovery perspective if the processing node fails. A proposed
-  enhancement to this scheme is to use another state, processing, to
-  indicate that the document has been retrieved and to update the
-  state to a processed state once processing is actually complete. For
-  the initial implementation we will use the processed state only.
+  document is considered \textit{processed} once retrieved from the
+  database. This may cause problems from a fault recovery perspective
+  if the processing node fails. A proposed enhancement to this scheme
+  is to use an intermediate state, \textit{processing}, to indicate
+  that the document has been retrieved and to update the state to
+  \textit{processed} once processing is actually complete. For the
+  initial implementation we will use the \textit{processed} state
+  only.
 
@@
  Figures \ref{fig:indexclassdiagram} and
- \ref{fig:indexsequencediagram} shows the UML class and sequence
+ \ref{fig:indexsequencediagram} show UML class and sequence
  diagrams respectively for classes and methods involved in the
@@
 \item \texttt{List<Document> search (String)}: given a comma-separated
-  list of search terms, returns the list of documents in which
-  the canonical form of all search terms appear together. This method
+  list of search terms, returns the list of documents in which the
+  canonical form of all search terms appear together. This method
   First calls \texttt{List<int> KeywordDAO.canonize (String)}, to map
@@
   We are assuming that the query will be formed with the set of
-  keywords, which will be a subset of the global keyword set, with which
-  the parsing state machine is constructed.
-
-  Then it calls \texttt{List<Integer> IndexDAO.find (List<querykeywordIDs>)},
-  which will searches in INDEX DB for a list of Common DOC\_IDs
-  corresponding to given query keywords.
+  keywords, which will be a subset of the global keyword set, with
+  which the parsing state machine is constructed.
+
+  Then it calls \texttt{List<Integer> IndexDAO.find
+    (List<querykeywordIDs>)}, which will searches in INDEX DB for a
+  list of Common DOC\_IDs corresponding to given query keywords.
 
@@
 Figures \ref{fig:searcherclassdiagram} and
-\ref{fig:searchersequencediagram} shows the UML class and sequence
+\ref{fig:searchersequencediagram} show UML class and sequence
 diagrams respectively for classes and methods involved in the
@@
 Figures \ref{fig:statemachineclassdiagram} and
-\ref{fig:statemachinesequencediagram} shows the UML class and sequence
+\ref{fig:statemachinesequencediagram} show UML class and sequence
 diagrams respectively for classes and methods involved in the
@@
 Figures \ref{fig:parserclassdiagram} and
-\ref{fig:parsersequencediagram} shows the UML class and sequence
+\ref{fig:parsersequencediagram} show UML class and sequence
 diagrams respectively for classes and methods involved in the
@@
 \label{sec:experimentresults}
-The whole idea of the project belongs in being able to reduce the time
-to build the index. As our system is building the inverted index by
-parsing all the post into the state machine, we can expect that the
-complexity will be proportional to the whole number of characters of
-all the post we want to build the index for. As a result changing the
-number of Posts, the number of characters inside of each post should
-have a big impact in our time complexity. Besides, we should also have
-a time needed at the very beginning of the running to build the state
-machine. We can expect that the building of the state machine needs a
-time proportional to the sum of the size of the keywords.
-
-The system proposed in this paper doesn’t take care of various
-compression techniques for storing the index. To do that, many
-techniques have been explored in the literature. We will just
-highlight the different sizes of our indexes according to different
-parameters.
-
-Finally we will do a time comparison between our way of building the
-inverted index and the naive way.  All the time complexities tests are
+The primary goal of this project is to be able to reduce the time
+to build a document index. As our system is building the inverted index by
+parsing all documents of the collection, we can expect the
+time complexity of the task to be proportional to the sum total of
+characters in the collection. As a result, changing the  
+number of documents or the average size of each document should
+have a big impact in our time complexity. In addition, our
+implementation will incur the cost of building the Aho-Corasick state
+machines used to scan the documents. We can expect the time to build a
+state machine to be proportional to the sum of the size of the
+keywords.
+
+Finally we will do a time comparison between our indexer
+implementation and a na\"{i}ve implementation. All the time
+complexities tests are  
 done on a single computer and it is important to understand that all
 these processes can be parallelized. A future work would be to install
-the system on multiple computers and explore the gain in time by
-running the code simultaneously on these computers.
+our indexer on multiple computers and explore the performance gain of
+running parallel instances simultaneously on all these computers.
 
 All tests were run on a single Intel Core 2 Duo 2GHz
-processor, 1GB RAM, running a 32 bit Linux 2.6.37 kernel. The Java
-Virtual Machine used was version 1.6.0-24. The database used was MySQL
-x.x, running on ....
+processor, 1GB RAM, running a 64-bit Linux 2.6.35 SMP kernel. The Java
+Virtual Machine used was version 1.6.0-24. The database used was
+MySQL, running on the same node as the indexer.
 
@@
 To be able to study the impact of different number of keywords, we
-built a keywords database with 10.000 most popular English
-words. These 10.000 words come from:
-http://www.mit.edu/~ecprice/wordlist.10000 All the experiments in this
-part are done with 50.000 posts randomly downloaded from internet. The
-idea is to get some heterogeneous length and type of post such that it
-is realistic.  The experiment shown here consist on building the index
-for different number of keywords from the 10.000 keywords database and
+built a keywords database with 10,000 most popular English
+words \cite{wordlist}. All the experiments in this
+part are done with 50,000 posts randomly downloaded from the Internet. The
+idea is to get a good distribution of documents with respect to length
+and content. The experiment shown here consist of building the index
+for different numbers of keywords from the 10,000 keyword database and
 explore the time and space complexity for the building of the state
-machine and the building of the index once the state machine is built.
+machine and the building of the index using the state machine.
 
@@
     \caption{Time to build Aho-Corasick state machine as a function of
-      the number of keywords. Time taken is shown in miliseconds on the
+      the number of keywords. Time taken is shown in milliseconds on the
       y-axis and the number of keywords on the x-axis}
@@
   to build the index increases logarithmically with the number of
-  keywords. This is likely due to the fac that as the number of keywords
+  keywords. This is likely due to the fact that as the number of keywords
   increases, so does the size of the index and the number of database
@@
     \caption{Time to build the inverted index as a function of the
-      number of keywords. Time taken is shown in miliseconds on the
+      number of keywords. Time taken is shown in milliseconds on the
       y-axis and the number of keywords on the x-axis}
@@
 means that our words have the same probability at the beginning or at
-the end of our 12,000 words dictionnary. This result is reflected in
+the end of our 12,000 word dictionary. This result is reflected in
 figure \ref{fig:numkeywordsspacecomplexbuildindex}.
@@
 To be able to study the impact of different number of Posts, we built
-a posts database with 400.000 posts randomly downloaded from
-internet. Then the keywords used to build the index are simply the
-first 5.000 from the previous keywords database.
+a posts database with 400,000 posts randomly downloaded from
+the Internet. Then the keywords used to build the index are simply the
+first 5,000 from the previous keywords database.
 
@@
     \caption{Time to build the inverted index as a function of the
-      size of the corpus. Time taken is shown in miliseconds on the
+      size of the corpus. Time taken is shown in milliseconds on the
       y-axis and the number of documents in the collection on the
@@
 \subsubsection{Space Complexity}
-As expected, the space used to sptore the index in the database is
+As expected, the space used to store the index in the database is
 proportional to the size of the corpus. This result is shown in figure
@@
 aspect and study the impact of the length of the post on the time to
-build the state machine. To do that we built a 5.000 posts database
-with posts that are all composed of 20.000 characters. Then, we ran
+build the state machine. To do that we built a 5,000 posts database
+with posts that are all composed of 20,000 characters. Then, we ran
 our system on different length of substring of these posts.
@@
   construction time is several orders of magnitude smaller than index
-  construction and negligible by comnparison. Our state machine is
+  construction and negligible by comparison. Our state machine is
   built in 770 ms on average for 1000 keywords with a standard
@@
     \caption{Time to build Aho-Corasick state machine as a function of
-      the average document size. Time taken is shown in miliseconds on the
+      the average document size. Time taken is shown in milliseconds on the
       y-axis and the average document size in characters on the x-axis}
@@
     \caption{Time to build an inverted index as a function of the
-      average document size. Time taken is shown in miliseconds on the
+      average document size. Time taken is shown in milliseconds on the
       y-axis and the average document size in characters on the x-axis} 
@@
 This part focuses on the improvement that our system with respect to
-time complexity in comparison with a na\"{i}e index builder. The na\"{i}ve
+time complexity in comparison with a na\"{i}ve index builder. The na\"{i}ve
 indexer builder has the same input as our system i.e. a list of
@@
   \caption{Time to build an inverted index as a function of the
-      average document size. Time taken is shown in miliseconds on the
+      average document size. Time taken is shown in milliseconds on the
       y-axis and the average document size in characters on the
@@
   \caption{Time to build an inverted index as a function of the
-      number of keywords. Time taken is shown in miliseconds on the
+      number of keywords. Time taken is shown in milliseconds on the
       y-axis and the number of keywords on the x-axis}
@@
   \caption{Time to build an inverted index as a function of the
-      number of keywords. Time taken is shown in miliseconds on the
+      number of keywords. Time taken is shown in milliseconds on the
       y-axis and the number of keywords on the logarithmic x-axis}
@@
       number of documents being indexed. Time taken is shown in
-      miliseconds on the y-axis and the number of documents on the
+      milliseconds on the y-axis and the number of documents on the
       x-axis} 
@@
       number of documents being indexed. Time taken is shown in
-      miliseconds on the y-axis and the number of documents on the
+      milliseconds on the y-axis and the number of documents on the
       logarithmic x-axis} 

### 2011-09-07 report/report.tex
Submitted
Added parallelization diagram

@@
 
-\begin{figure}
+\begin{figure}[h!]
   \begin{center}
@@
 implementation and a na\"{i}ve implementation. All the time
-complexities tests are  
-done on a single computer and it is important to understand that all
-these processes can be parallelized. A future work would be to install
-our indexer on multiple computers and explore the performance gain of
-running parallel instances simultaneously on all these computers.
-
-All tests were run on a single Intel Core 2 Duo 2GHz
-processor, 1GB RAM, running a 64-bit Linux 2.6.35 SMP kernel. The Java
-Virtual Machine used was version 1.6.0-24. The database used was
-MySQL, running on the same node as the indexer.
+complexities tests are done on a single computer and it is important
+to understand that all these processes can be parallelized. In section
+\ref{parallelization} we install our indexer on multiple computers and
+explore the performance gain of running parallel instances
+simultaneously.
+
+All tests were run on a single Intel Core 2 Duo 2GHz processor, 1GB
+RAM, running a 64-bit Linux 2.6.35 SMP kernel. The Java Virtual
+Machine used was version 1.6.0-24. The database used was MySQL,
+running on the same node as the indexer.
 
@@
 
-  \begin{figure}
+  \begin{figure}[h]
     \begin{center}
@@
 
-  \begin{figure}
+  \begin{figure}[h!]
     \begin{center}
@@
 As expected, the two implementations exhibit a linear response to the
-number of keywords used. In the na\"{i}ve implmementation:
+number of keywords used. In the na\"{i}ve implemementation:
 
@@
 
-The experiment results are shown in figure \ref{fig:naivedocumentsize}
-and on a logarithmic scale in figure \ref{fig:naivedocumentsizelog}.
-
-\begin{figure}
+The experiment results are shown in figure \ref{fig:naivenumkeywords}
+and on a logarithmic scale in figure \ref{fig:naivenumkeywordslog}.
+
+\begin{figure}[p]
   \begin{center}
@@
 
-\begin{figure}
+\begin{figure}[p]
   \begin{center}
@@
   \label{fig:naivesizecorpuslog}
+\end{figure}
+
+%----------------------------------------
+% Parallelization
+%----------------------------------------
+\subsection{Parallelization}
+\label{parallelization}
+
+\begin{figure}[p]
+  \begin{center}
+    \includegraphics[width=\textwidth,height=!]{paralleldeploymentmodel}
+  \end{center}
+  \caption{Deployment model of parallel indexer} 
+  \label{fig:paralleldeploymentmodel}
 \end{figure}

### 2011-09-08 report/report.tex
Added parallelization graph
First draft complete

@@
 \label{parallelization}
-
-\begin{figure}[p]
+Given the modular architecture of our indexer, it is possible to
+deploy multiple instances of the program to work concurrently on the same
+indexing task. The instances interact via the database which
+synchronizes their operations on the corpus and the index itself. To
+investigate the effect of parallelization on the indexer operation, the
+indexer was deployed as follows:
+
+\begin{itemize}
+\item Single database server storing documents, keywords, and the
+  resulting index.
+\item Four indexer nodes each running a single instance of the indexer
+  program. 
+\end{itemize}
+
+This deployment as well as a description of the sequence of operation
+are shown in figure \ref{fig:paralleldeploymentmodel}.
+
+\begin{figure}[h]
   \begin{center}
@@
 \end{figure}
+
+In this experiment a corpus of 100,00 documents and 1000 keywords were
+used, each document 5000 words in length. Tests were run using various
+node combinations and the time taken to index the corpus was measured and
+normalized to a throughput metric (documents/second) for convenient
+comparison. Test results using the same number of nodes were averaged
+to account for differences in the physical capabilities and
+configuration of the worker nodes. Next, a series of tests was
+performed using a corpus of 800,000 documents to investigate the
+effect of corpus size on performance. The results of this experiment
+are shown in figure \ref{fig:parallelization} and described in detail
+in the following section.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=\textwidth,height=!]{parallelization}
+  \end{center}
+  \caption{Performance characteristics of parallel indexer
+    deployment. The \textit{ideal} curve shows the theoretical
+    performance curve based on linear combination of average
+    individual node performance over the corpus} 
+  \label{fig:parallelization}
+\end{figure}
+
+\subsubsection{Results}
+For the 100,000 document corpus, increasing the number of nodes
+increases throughput but with diminishing returns. As the database
+connection pool is saturated, this becomes the limiting factor of the
+architecture. If the number of nodes were increased further, one would
+expect the curve to plateau at some point. After this point, adding
+additional nodes would offer no improvement and may even decrease
+performance as the worker processes spend an increasing amount of time
+context switching and contending for locks. These factors could be
+mitigated up to a point by increasing batch sizes and tuning the
+database. However, the characteristics of the curve are ultimately
+unavoidable in the current architecture.
+
+For the 800,000 document corpus, the performance curve shows an even
+more severe degradation than the smaller corpus. Since we normalize
+the performance metrics for the size of the document collection, one
+would expect this curve to be more similar to that of the 100,000
+document collection. A possible explanation for the deviation is that
+for the larger collection, more documents were present in the
+database and the resulting index larger. This results in longer table
+scan times to retrieve and update records. This effect could be
+alleviated through the use of indices and other database optimization
+techniques, however, as for the smaller document
+collection, this degradation is ultimately unavoidable.
 
@@
 effect of parallelization in reducing the time to index a document
-collection. Due to time constraints, this was not done but the
-resulting implementation lays a strong foundation for further research
-in this area. Other avenues of interest include, but are not limited
+collection. We successfully demonstrated that performance gains could
+be achieved for the indexing task through parallelization as well as
+the fact that our architecture offers diminishing returns as the
+number of nodes is increased and as the corpus becomes larger.
+
+ Other avenues of interest include, but are not limited
 to:
@@
       reduce the index size in the database.
+
+      \item Conduct further parallelization experiments to determine
+        the performance characteristics of our indexer over a wider
+        range of operating conditions.
 

### 2011-10-11 report/article.tex
Started article template

--- 2011-09-08 report/report.tex
@@
 %
-% $Id$ 
+% $Id: report.tex 52 2011-10-03 17:08:43Z justinkamerman $ 
 %
-% $LastChangedDate$ 
+% $LastChangedDate: 2011-10-03 14:08:43 -0300 (Mon, 03 Oct 2011) $ 
 % 
-% $LastChangedBy$
+% $LastChangedBy: justinkamerman $
 %
 
-\documentclass[10pt]{report}
+\documentclass[10pt]{article}
 \usepackage{graphicx}
@@
 \usepackage{algorithmic}
-\usepackage{setspace}                   
-\onehalfspacing
-% numbering down to subsubsection
-\setcounter{secnumdepth}{3}
 
-\title{CS6999 Web Data and Text Mining:\\Parallel Text Indexer Project Report}
-\author{Damien DuBois\\Justin Kamerman 3335272\\Ramanpreet Singh 3395913}
+\title{Document Indexing Aho-Corasick State Machines}
+\author{Damien DuBois, Justin Kamerman, Ramanpreet Singh}
 \date{\today}
@@
 \maketitle
-% No chapter numbers
-\renewcommand*\thesection{\arabic{section}}
+
+%----------------------------------------
+% Abstract
+%----------------------------------------
+\begin{abstract}
+The Aho-Corasick algorithm was originally proposed as a bibliographic
+search mechanism, efficient enough to preclude the construction and
+maintenance of a search index. As the size of document collections
+have grown since the development of the algorithm, document
+collections have grown to the point where cost of rescanning a corpus
+for every search has become prohibitive and indexing, in some form or
+another, is an essential optimization for any modern information
+retrieval system. In this paper we demonstrate the use of
+Aho-Corasick, not as an alternative to indexing but as a lexical
+scanning tool employed in the construction of an inverted index. We
+provide analytic predictions of our indexing algorithm and an
+empirical evaluation of our inverted index implementation is conducted
+and the results analyzed. Consistent with our analysis, the hybrid
+approach yields significant improvement over a naive
+implementation. In addition, the modular architecture of our indexer
+allowed the indexing task to be parallelized over multiple physical
+nodes. The results of parallelization are presented and offer
+interesting insights towards further horizontal scaling.
+\\\\
+{\bf Keywords:} inverted index, Aho-Corasick state machine, finite
+automata, pattern matching, parsing
+\end{abstract}
 
@@
 \section{Introduction}
-Fundamental to the real-world application of \textit{Information
-  Retrieval} techniques is the ability to objectively measure the
-quality of a particular algorithm or system and compare it to
-others. Often, information retrieval algorithms lack an analytical
-model that would allow one to make formal statements as to the
-performance characteristics of a particular system. In addition, as with
-machine learning and classification problems, using
-a scalar value to represent the performance of a system as a whole is
-misleading and does not account for the variety of operating
-conditions under which the system may be expected to perform.
+Information Retrieval is a fast becoming the dominant form of
+information access, overtaking traditional relational database style
+searching\cite{RefWorks:109}. The basis of most document retrieval systems is the
+term-document incidence matrix. This matrix is typically sparse and
+more efficiently represented as an inverted-index which maps terms to
+the parts of a document where they occur.
 
-In measuring the performance of an information retrieval system, the
-following factors constitute a set of operating conditions which
-qualify the result:
-
-\begin{itemize}
-  \item Accuracy of the retrieval result with respect to the
-    query. Common measures in this respect are \textit{Precision} and
-    \textit{Recall}. Search techniques employing latent semantics of
-    document collections are subtly more difficult to quantify, though
-    measurable nonetheless.
-
-    \item Speed with which queries are answered. This metric is a
-      function of the number of search terms and the size of the
-      corpus.
-
-    \item Size of the document collection and, if not static, the rate
-      at which new documents are being added and/or removed.
-
-    \item The computational cost of the technique. Multiple
-      activities may contribute to this aspect: searching, indexing,
-      preprocessing, and post processing. It is important to consider
-      whether the technique allows indices or intermediate
-      approximations of the document collection to be altered
-      incrementally or whether they must be regenerated to incorporate
-      new documents. Also significant is whether the technique can be
-      parallelized and thereby benefit from concurrency.
-
-    \item The storage requirements of the technique. How do the
-      storage requirements evolve with the size of the collection
-      and/or degree of concurrency ?
-\end{itemize}
-
-To drive information retrieval research in one direction or another,
-it would be useful to have a physical implementation
-framework within which one could empirically explore the operating
-characteristics of a particular algorithm and/or technique along the lines
-described above. Such a system would allow researchers to investigate
-how the system responds to variation in any one of these operating
-parameters. The aim of this project is to create such an
-information retrieval \textit{test-bed}.
-
-The initial reference implementation of this information retrieval
-test-bed is based on a simple \textit{inverted index} but has been
-architected such that the specific algorithm could easily be changed
-without significant rework. This paper presents a survey of existing
-literature in support of our test-bed concept, as well as the specific
-algorithms employed in the reference implementation. In
-addition, a detailed architectural overview of the system is
-presented, giving insight into how the system would be modified for
-alternative algorithms. Lastly, an evaluation of our inverted index
-reference implementation is conducted and the results analyzed.
-
-%----------------------------------------
-% Literature Review
-%----------------------------------------
-\section{Literature Review}
-\label{sec:literaturereview}
-
-\subsection{Indexing}
-\textit{Information Retrieval} is a fast becoming the dominant form of
-information access, overtaking traditional relational database style
-searching \cite{RefWorks:109}. The basis of most document retrieval
-systems is the \textit{term-document} incidence matrix. This matrix is
-typically sparse and more efficiently represented as an
-\textit{inverted-index} which maps terms to the parts of a document
-where they occur.
-
-In order to construct an \textit{inverted index}, documents must be
-scanned to determine \termit{term frequencies}. The Aho-Corasick
-string matching algorithm\cite{RefWorks:103} is a simple and efficient
-text scanning algorithm. The algorithm constructs a finite
-state machine to scan for a given set of keywords. It is, in effect, a
-reduced grammar regular expression parser of the type described in
-\cite{RefWorks:111}. The algorithm is simple and efficient,
-construction time being proportional to the sum of the length of the
-keyword set, and the number of state transitions required to scan a
-document is independent of the number of the size of the keyword
-set.
-
-\subsection{Evaluation}
-In the case of machine learning and classification systems, the
-specification of a particular set of misclassification costs or false
-and positive misclassification rates define a particular
-\textit{operating point} of the system.
-
-This concept applies equally to \textit{information retrieval} systems
-whereby it is misleading to represent the operating characteristics of
-a particular system with a scalar metric. In addition, the evaluation
-of information retrieval systems suffers from the the inherent
-subjectivity with respect to \textit{relevance}
-\cite{RefWorks:72}. Non-scalar evaluation methods try to characterize
-the performance of a system over a range of operating
-conditions. Although this makes direct performance comparisons
-ambiguous, we are more likely to be able to select the algorithm which
-will perform best in the real world by applying domain expertise to
-the results of a non-scalar evaluation \cite{RefWorks:32}.
-
-\subsection{Parallelization}
-Once an objective measure has been established, the performance of an
-information retrieval system can be improved iteratively by making
-various optimizations. These optimizations may take the form of
-algorithm enhancements, implementation efficiencies, or
-parallelization.
-
-Algorithm enhancements usually represent the forefront of research in
-the field. Work to validate a new concept or algorithm is usually
-performed in a very controlled environment and on relatively small
-document collections. As the field matures, improvements are harder
-fought and often involve increasingly complex algorithms and corpus
-approximation methods into which it becomes more difficult to
-\textit{fold-in} new documents. In addressing the considerable
-computation cost of re-calculating \textit{Latent Semantic Indexing}
-(LSI) \cite{RefWorks:102} approximations, \cite{RefWorks:101}
-introduced new update methods which saw improved average precisions
-but were forced to conclude that the improvements were at a
-computational premium over previous methods. A semi-discrete matrix
-decomposition for LSI, \textit{SDD}, performs as well as the original
-SVD-based LSI method for equal query times and uses less storage
-\cite{RefWorks:99}. The decomposition time of the new method, however,
-is significantly larger than the original.
-
-Enhancement of implementation efficiency in \cite{RefWorks:96} saw a
-six-fold timing improvement in \textit{LSI} searches on large document
-collections. The same study achieved nearly 180 times improvement
-through the use of parallelization. The study made no algorithmic
-improvements and the result is consistent with increased use of
-parallelization by large commercial systems to cope with the ever
-increasing bodies of data that must be processed.
-
-\textit{Map-Reduce}, popularized by Google, is emerging as an
-important programming model for large-scale data-parallel applications
-such as web indexing \cite{RefWorks:113}. Fundamentally, Map-Reduce
-breaks up a computation into smaller tasks which are made to run in
-parallel on different physical machines. The \textit{map} and
-\textit{reduce} phases are inspired by the functions of the same name
-used in \textit{functional programming}. The model scales well to large
-clusters mostly because of the decoupled nature of the tasks
-themselves. An initial \textit{map} phase breaks the data into
-  separate chunks and assigns them to a task. The \textit{reduce}
-  phase assembles the results of the map tasks to produce the final
-  result.
-
-\textit{Hadoop} is an open source implementation of the MapReduce
-model, developed mostly by Yahoo, and now maintained by the Apache
-Software Foundation. Hadoop can be configured to automatically create
-speculative copies of slow running map tasks
-(stragglers). This facility is intended as a pre-emptive failure
-recovery mechanism as well as a means of optimizing overall
-throughput. The metrics used to identify stragglers as well as the
-algorithm for determining the target node on which speculative copies
-of a task should be launched, is examined in \cite{RefWorks:128}. The
-paper introduces an alternative scheduling algorithm, \textit{LATE},
-to address excessive task speculation issues with the native Hadoop
-scheduler in a heterogeneous hardware environment. Experiments
-comparing the native Hadoop scheduler, a non-speculative scheduling
-strategy, and LATE were conducted using hosted virtual machines (VMs)
-in the Amazon EC2 environment as well as a dedicated local
-cluster. Variance in node performance was introduced by varying the
-distribution of VMs across physical machines and also introducing
-network and disk transfer loads. Results show speculative execution
-(native Hadoop and LATE) outperforming non-speculative execution and
-LATE outperforming native Hadoop in most cases.
+In order to construct an inverted index, documents must be scanned to
+determine term frequencies. The Aho-Corasick string matching
+algorithm\cite{RefWorks:103} is a simple and efficient text scanning algorithm. The
+algorithm constructs a finite state machine to scan for a given set of
+keywords. It is, in effect, a reduced grammar regular expression
+parser of the type described in \cite{RefWorks:111}. The algorithm is simple and
+efficient, construction time being proportional to the sum of the
+length of the keyword set, and the number of state transitions
+required to scan a document is independent of the number of the size
+of the keyword set. Aho-Corasick was originally proposed as a
+bibliographic search mechanism, efficient enough to preclude the
+construction and maintenance of a search index. As the size of
+document collections have grown since the publishing of the
+Aho-Corasick algorithm, document indexing is essential for efficient
+document search. In this paper we demonstrate the use of Aho-Corasick,
+not as an alternative to indexing but as a lexical scanning tool
+employed in the construction of an inverted index. We provide analytic
+predictions of our implementation against a naive algorithm and
+finally, an empirical evaluation of our inverted index implementation
+is conducted and the results analyzed.
 
@@
 In addition, the reference implementation will be compared against a
-na\"{i}ve implementation of the inverted indexer. The na\"{i}ve
-implementation provided a baseline for further evaluation.
+naïve implementation of the inverted indexer. The naïve implementation
+provided a baseline for further evaluation.
 
@@
   \label{fig:deploymentmodel}
-\end{figure}
-
-A guiding architectural principal of the indexer design is to separate the implementation
-into three logical layers:
-\begin{itemize}
-\item \textbf{Software Layer:} implements parsing, indexing and search
-  algorithms.
-\item \textbf{Data Access Layer:} implements an
-  \textit{object-relational mapping}, insulating algorithm
-  implementation from the persistence details. This layer implements
-  classes which encapsulate persistent entities within the system.
-\item \textbf{Database:} the persistent store for the document
-  collection and inverted index.
-\end{itemize}
-
-The implementation classes and their relationships are represented in
-a UML class diagram in figure \ref{fig:overallclassdiagram}. Figure
-\ref{fig:overallsequencediagram} is a UML sequence diagram showing how
-the layers interact at a high level.  The implementation of each of
-these layers is described in more detail in section
-\ref{sec:implementation}.
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[angle=90,width=\textwidth,height=!]{overallclassdiagram}
-  \end{center}
-  \caption{Overall UML class diagram design}
-  \label{fig:overallclassdiagram}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{overallsequencediagram}
-  \end{center}
-  \caption{Overall UML sequence diagram design}
-  \label{fig:overallsequencediagram}
-\end{figure}
-
-%----------------------------------------
-% Implementation 
-%----------------------------------------
-\section{Implementation}
-\label{sec:implementation}
-The individual worker processes of the indexer system are implemented as Java
-programs. The programs make use of the following thirdparty libraries:
-
-\begin{itemize}
-  \item \textbf{DBPool:} a database connection pooling library, used
-    to manage connections to relational databases.
-  \item \textbf{Apache Commons Logging:} a generic logging API used by
-    DBPool.
-  \item \textbf{Apache Commons CLI:} a library for parsing command
-    line arguments passed to Java programs.
-  \item \textbf{MySQL JDBC driver:} the MySQL Java client driver.
-\end{itemize}
-
-The indexer system maintains an \textit{inverted index} of an evolving
-corpus and services concurrent search requests against the index. The
-current implementation is based on boolean retrieval methods described
-in \cite{RefWorks:109}. An Aho-Corasick state machine
-\cite{RefWorks:103} is used to scan documents for a predefined set of
-keywords, and construct a term index for each document (delta
-index). The term occurrences for each document are used to augment an
-existing index of documents already processed. The document
-collection, index, and keyword list are maintained in a single MySQL
-database instance. It is through this database that the parallel
-indexer instances are initialized and synchronized. This configuration
-is operationally simple and easy to implement but the database has the
-potential to become a throttle point as the system is scaled to larger
-and more rapidly evolving document collections, and is required to
-service a higher rate of queries. In order to scale the current
-implementation to large document collections, a more efficient
-mechanism would be required in this respect.
-
-The following subsections detail the implementation of the
-architectural layers identified in section \ref{sec:design}.
-
-% Data Access Layer
-\subsection{Data Access Layer}
-\label{sec:dataaccesslayer}
-Access to persistent storage is via objects of the data access
-layer. Confining data access to a logical grouping of classes
-insulates the other application components from changes in database
-schema. The following subsections detail the individual data entities
-handled by the data access later.
-
-% Document
-\subsubsection{Document}
-\label{sec:document}
-The \texttt{DocumentDAO} singleton class provides methods for retrieving
-\texttt{Document} objects from the document database:
-
-\begin{itemize}
-\item \texttt{List<Document> getDocuments (int count)}: retrieve a
-  specified number of unprocessed documents from the database. The
-  documents retrieved will be atomically marked as processed so that
-  processing is not duplicated on other nodes. In this scheme, a
-  document is considered \textit{processed} once retrieved from the
-  database. This may cause problems from a fault recovery perspective
-  if the processing node fails. A proposed enhancement to this scheme
-  is to use an intermediate state, \textit{processing}, to indicate
-  that the document has been retrieved and to update the state to
-  \textit{processed} once processing is actually complete. For the
-  initial implementation we will use the \textit{processed} state
-  only.
-
-\item \texttt{Document getDocumentById (int id)}: retrieve a specific
-  document from the database, referenced by its ID.
-
-\item \texttt{int getID ()}: accessor method.
-\item \texttt{String getContent ()}: accessor method.
-\item \texttt{String getAuthor ()}: accessor method.
-\item \texttt{String getTitle ()}: accessor method
-\item \texttt{String getLink()}: accessor method.
-\end{itemize}
-
-The \texttt{Document} class encapsulates the BLOG database table, with the
-addition of a state field that tracks whether a document has been
-processed or not. Figures \ref{fig:documentclassdiagram} and
-\ref{fig:documentsequencediagram} show class and sequence diagrams
-respectively for non-trivial classes and methods involved in the
-implementation of the Document data access layer.
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{documentclassdiagram}
-  \end{center}
-  \caption{Document class diagram}
-  \label{fig:documentclassdiagram}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{documentsequencediagram}
-  \end{center}
-  \caption{Document sequence diagram}
-  \label{fig:documentsequencediagram}
-\end{figure}
-
-% Keyword
-\subsubsection{Keyword}
-The \texttt{KeywordDAO} singleton class provides methods for
-retrieving \texttt{Keyword} objects from the document
-database. Because the keyword set is relatively static and accessed
-often, the full set of keywords will be loaded from the database and
-cached when the class initializes. The class also provides methods for
-converting search terms to their canonical form.
-
-\begin{itemize}
-\item \texttt{List<String> getKeywords ()}: this method
-  will returns a list of all \texttt{Keyword} objects defined in the
-  keyword database.
-
-\item \texttt{List<Integer> canonize (String query)}: return the list
-  of IDs corresponding to a comma-separated list of query terms.
-
-\item \texttt{Keyword getKeywordById (int id)}: return the
-  \texttt{Keyword} object with the given identifier.
-\end{itemize}
-
-The \texttt{Keyword} field encapsulates a canonical search term and a
-list of its synonyms. Figures \ref{fig:keywordclassdiagram} and
-\ref{fig:keywordsequencediagram} show UML class and sequence diagrams
-respectively for non-trivial classes and methods involved in the
-implementation of the Keyword data access layer.
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{keywordclassdiagram}
-  \end{center}
-  \caption{Keyword class diagram}
-  \label{fig:keywordclassdiagram}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{keywordsequencediagram}
-  \end{center}
-  \caption{Keyword sequence diagram}
-  \label{fig:keywordsequencediagram}
-\end{figure}
-
-% Index
-\subsubsection{Index}
-The \texttt{IndexDAO} class provides methods for updating and
-searching the inverted index table stored in the database. There are
-two methods defined in this class:
-
-\begin{itemize}
-\item \texttt{void UpdateIndex (List<TempIndex> input)}: this method
-  takes input of type List<TempIndex> (Note: Temporary index is a user
-  defined data class, which contains Document ID and List of <keyword
-  ids> as its attributes) and generates an SQL query which updates the
-  INDEX table in the Database.
-
-\item \texttt{List<Integer> Find (List<Integer> QueryKeywordIds)}:
-  this method takes a list of \texttt{Integers} (which corresponds to
-  mapped \texttt{String} query into keyword ids) as input and searches
-  the INDEX table in DB to find all the document ids which are common
-  between for the given query keywords. 
-\end{itemize}
-
- Figures \ref{fig:indexclassdiagram} and
- \ref{fig:indexsequencediagram} show UML class and sequence
- diagrams respectively for classes and methods involved in the
- implementation of the Index data access layer.
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{indexclassdiagram}
-  \end{center}
-  \caption{Index class diagram}
-  \label{fig:indexclassdiagram}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{indexsequencediagram}
-  \end{center}
-  \caption{Index sequence diagram}
-  \label{fig:indexsequencediagram}
-\end{figure}
-
-% Software Layer
-\subsection{Software Layer}
-
-% Searcher
-\subsubsection{Searcher}
-This class is designed to perform searches for a query entered by the
-user. It contains a method named \texttt{search ()} which is responsible for
-finding all the documents which contained all query keywords.
-
-\begin{itemize}
-\item \texttt{List<Document> search (String)}: given a comma-separated
-  list of search terms, returns the list of documents in which the
-  canonical form of all search terms appear together. This method
-  First calls \texttt{List<int> KeywordDAO.canonize (String)}, to map
-  string query keywords into corresponding list of keyword IDs.  Note:
-  We are assuming that the query will be formed with the set of
-  keywords, which will be a subset of the global keyword set, with
-  which the parsing state machine is constructed.
-
-  Then it calls \texttt{List<Integer> IndexDAO.find
-    (List<querykeywordIDs>)}, which will searches in INDEX DB for a
-  list of Common DOC\_IDs corresponding to given query keywords.
-
-  In the end, it calls \texttt{List<Document>
-    DocumentDAO.getDocumentById (DOC\_ID)}. This will map DOC\_ID into
-  a \texttt{Document} from the BLOG table in the database. 
-\end{itemize}
-
-Figures \ref{fig:searcherclassdiagram} and
-\ref{fig:searchersequencediagram} show UML class and sequence
-diagrams respectively for classes and methods involved in the
-implementation of the \texttt{Searcher} class.
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{searcherclassdiagram}
-  \end{center}
-  \caption{Searcher class diagram}
-  \label{fig:searcherclassdiagram}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{searchersequencediagram}
-  \end{center}
-  \caption{Searcher sequence diagram}
-  \label{fig:searchersequencediagram}
-\end{figure}
-
-% State Machine
-\subsubsection{State Machine}
-This class is designed to build the state machine from the list of
-keywords returned by the \texttt{KeywordDAO.getKeywords()} method. This class has a static
-attribute \texttt{initialState} that is the first state of the state machine
-built by the function \texttt{createStateMachine} (initializes with the value
-null). The first state gives us the whole state machine thanks to the
-architecture of a state. Once the state machine is built, the function
-\texttt{getStateMachine} is used to access the initial state stored in the
-class.
-
-Figures \ref{fig:statemachineclassdiagram} and
-\ref{fig:statemachinesequencediagram} show UML class and sequence
-diagrams respectively for classes and methods involved in the
-implementation of the \texttt{StateMachineBuilder}.
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{statemachineclassdiagram}
-  \end{center}
-  \caption{State machine class diagram}
-  \label{fig:statemachineclassdiagram}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{statemachinesequencediagram}
-  \end{center}
-  \caption{State machine sequence diagram}
-  \label{fig:statemachinesequencediagram}
-\end{figure}
-
-% Parser
-\subsubsection{Parser}
-The \texttt{Parser} implements the document parsing algorithm within
-the state machine. By default, it uses the state machine contained
-in the static attribute of \texttt{StateMachineBuilder}. To parse a
-document we need to call the \texttt{parse ()} function inside this
-class, passing it the \texttt{Document} class encapsulating the actual
-text we want to parse. This class returns a \texttt{TemporaryIndex}
-object which encapsulates a reference to the document parsed and the
-keywords found to occur within the document.
-
-Figures \ref{fig:parserclassdiagram} and
-\ref{fig:parsersequencediagram} show UML class and sequence
-diagrams respectively for classes and methods involved in the
-implementation of the \texttt{Parser}.
-
-\begin{figure}[h!]
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{parserclassdiagram}
-  \end{center}
-  \caption{Parser class diagram}
-  \label{fig:parserclassdiagram}
-\end{figure}
-
-\begin{figure}[h!]
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{parsersequencediagram}
-  \end{center}
-  \caption{Parser sequence diagram}
-  \label{fig:parsersequencediagram}
 \end{figure}
@@
 
-In this experiment a corpus of 100,00 documents and 1000 keywords were
+In this experiment a corpus of 100,000 documents and 1000 keywords were
 used, each document 5000 words in length. Tests were run using various
@@
 
-%--------------------------------------------------
 \end{document}

### 2011-10-20 report/article.tex
Paper down to 11 pages - sending for review

@@
 \usepackage{algorithmic}
+\usepackage{wrapfig}
+\usepackage{caption3} % load caption package kernel first
+\DeclareCaptionOption{parskip}[]{} % disable "parskip" caption option
+\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
 
-\title{Document Indexing Aho-Corasick State Machines}
+\title{Document Indexing Using Aho-Corasick State Machines}
 \author{Damien DuBois, Justin Kamerman, Ramanpreet Singh}
@@
 Aho-Corasick, not as an alternative to indexing but as a lexical
-scanning tool employed in the construction of an inverted index. We
-provide analytic predictions of our indexing algorithm and an
-empirical evaluation of our inverted index implementation is conducted
-and the results analyzed. Consistent with our analysis, the hybrid
-approach yields significant improvement over a naive
-implementation. In addition, the modular architecture of our indexer
+scanning tool employed in the construction of an inverted
+index. Consistent with analytical predictions, an empirical evaluation
+of our hybrid indexer, shows significant improvement over a
+na\"{\i} implementation. In addition, the modular architecture of our indexer
 allowed the indexing task to be parallelized over multiple physical
@@
 \section{Introduction}
-Information Retrieval is a fast becoming the dominant form of
-information access, overtaking traditional relational database style
-searching\cite{RefWorks:109}. The basis of most document retrieval systems is the
+The basis of most document retrieval systems is the
 term-document incidence matrix. This matrix is typically sparse and
@@
 %----------------------------------------
-% Research Plan
-%----------------------------------------
-\section{Research Plan}
-\label{sec:researchplan}
-The primary goals of this research project is to construct an
-information retrieval framework within which specific implementations
-can be evaluated under various operating conditions. To initiate the
-synthesis of this general framework, a reference implementation will
-be built, based on the notion of a simple inverted index.
-
-The software architecture of the reference implementation will be
-carefully documented and follow established design patterns in order
-to facilitate the future construction of alternative document search
-schemes within the same framework. The framework will be designed to
-be executed in a distributed deployment such that it may be used to
-evaluate the effect of parallelization on a particular algorithm.
-
-The performance of the reference implementation will be tested with
-respect to the following operating parameters and compared to analytic
-expectations:
-
-\begin{itemize}
-  \item Number of index keywords
-  \item Number of documents in collection
-  \item Average size of documents
-\end{itemize}
-
-In addition, the reference implementation will be compared against a
-naïve implementation of the inverted indexer. The naïve implementation
-provided a baseline for further evaluation.
-
-%----------------------------------------
 % Design
 %----------------------------------------
-\section{Design}
-\label{sec:design}
-A black-box functional view of the indexing system is show in figure
-\ref{fig:blackbox}. The system processes a continuous incoming flow of
-documents and distribute them to parallel indexing threads running on
-physically separate, heterogeneous nodes.  Document queries are performed
-using an evolving search index. The operational balance is to
-timeously index new additions to the corpus while servicing concurrent
-search requests. A view as to how the components of the indexer system
-are deployed is show in figure \ref{fig:deploymentmodel}.
-
+\section{Implementation}
+\label{sec:implementation}
 As can be seen in figure \ref{fig:deploymentmodel}, the indexer
-processes are symmetrical and deployed on multiple physical nodes. The
+processes are symmetrical and can deployed on multiple physical nodes. The
 individual indexers do not interact directly with one another, making
@@
 
-\begin{figure}
+\begin{wrapfigure}{r}{0.70\textwidth}
   \begin{center}
-        \includegraphics[width=\textwidth,height=!]{blackbox}
-  \end{center}
-  \caption{A black-box view of the indexer}
-  \label{fig:blackbox}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-        \includegraphics[width=\textwidth,height=!]{deploymentmodel}
+        \includegraphics[width=0.70\textwidth,height=!]{deploymentmodel}
   \end{center}
@@
   \label{fig:deploymentmodel}
-\end{figure}
+\end{wrapfigure}
 
@@
 
-\subsubsection{Time Complexity}
-\begin{itemize}
-\item \textbf{State Machine Construction:}
 The time taken to construct the Aho-Corasick state machine was
 measured for different numbers of keywords. The results of this test
-  is shown in figure \ref{fig:numkeywordstimecomplexbuildsm} and construction time
-  can easily be considered linear with respect to the number of
-  keywords. This is consistent with \cite{RefWorks:103} which proves
-  that the state machine construction algorithm is linearly
+is shown in figure \ref{fig:numkeywordstimecomplexbuildsm} and
+construction time can easily be considered linear with respect to the
+number of keywords. This is consistent with \cite{RefWorks:103} which
+proves that the state machine construction algorithm is linearly
 proportional to the sum of the lengths of the keywords used to
@@
 
-  \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=\textwidth,height=!]{numkeywordstimecomplexbuildsm}
-    \end{center}
-    \caption{Time to build Aho-Corasick state machine as a function of
-      the number of keywords. Time taken is shown in milliseconds on the
-      y-axis and the number of keywords on the x-axis}
-    \label{fig:numkeywordstimecomplexbuildsm}
-  \end{figure}
-
-\item \textbf{Index Construction:} 
 The time taken for a state machines to build an inverted index of our
@@
 to build the index increases logarithmically with the number of
-  keywords. This is likely due to the fact that as the number of keywords
-  increases, so does the size of the index and the number of database
-  interacts required to store and update the index.
+keywords. This is likely due to the fact that as the number of
+keywords increases, so does the size of the index and the number of
+database interacts required to store and update the index.
 
-  \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=\textwidth,height=!]{numkeywordstimecomplexbuildindex}
-    \end{center}
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{numkeywordstimecomplexbuildsm}
+    \caption{Time to build Aho-Corasick state machine as a function of
+      the number of keywords. Time taken is shown in milliseconds on the
+      y-axis and the number of keywords on the x-axis}
+    \label{fig:numkeywordstimecomplexbuildsm}
+  \end{minipage}
+  \hspace{0.5cm}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{numkeywordstimecomplexbuildindex}
     \caption{Time to build the inverted index as a function of the
@@
     \label{fig:numkeywordstimecomplexbuildindex}
+  \end{minipage}
 \end{figure}
 
-\end{itemize}
-
-\subsubsection{Space Complexity}
-As we would have expected it, the space used by the index in the
-database is proportional to the number of words we are using. It also
-means that our words have the same probability at the beginning or at
-the end of our 12,000 word dictionary. This result is reflected in
-figure \ref{fig:numkeywordsspacecomplexbuildindex}.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\textwidth,height=!]{numkeywordsspacecomplexbuildindex}
-  \end{center}
-    \caption{Size of index as a function of the number of
-      keywords. The number of database rows used to store the index is
-      shown on the y-axis and the number of keywords on the x-axis} 
-    \label{fig:numkeywordsspacecomplexbuildindex}
-\end{figure}
+The space used by the index in the database is proportional to the
+number of words we are using. It also means that our words have the
+same probability at the beginning or at the end of our 12,000 word
+dictionary.
 
@@
 
-\subsubsection{Time Complexity}
 As we were expecting, the time to build the index with the state
@@
 
-\begin{figure}[h!]
-  \begin{center}
-    \includegraphics[width=\textwidth,height=!]{corpsizetimecomplexbuildindex}
-  \end{center}
+The space used to store the index in the database is
+proportional to the size of the corpus as shown in figure
+\ref{fig:corpsizespacecomplexbuildindex}.
+
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{corpsizetimecomplexbuildindex}
     \caption{Time to build the inverted index as a function of the
@@
     \label{fig:corpsizetimecomplexbuildindex}
-\end{figure}
-
-\subsubsection{Space Complexity}
-As expected, the space used to store the index in the database is
-proportional to the size of the corpus. This result is shown in figure
-\ref{fig:corpsizespacecomplexbuildindex}.
-
-\begin{figure}[h!]
-  \begin{center}
-    \includegraphics[width=\textwidth,height=!]{corpsizespacecomplexbuildindex}
-  \end{center}
+  \end{minipage}
+  \hspace{0.5cm}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{corpsizespacecomplexbuildindex}
     \caption{Size of index as a function of the size of the
@@
     \label{fig:corpsizespacecomplexbuildindex}
+  \end{minipage}
 \end{figure}
@@
 
-\subsubsection{Time Complexity}
-\begin{itemize}
-\item \textbf{State Machine Construction:} 
-  As expected, state machine construction time
-  does not depend on the length of the documents. Nonetheless it is
-  interesting to compare figures \ref{fig:docsizetimecomplexbuildsm}
-  and \ref{fig:docsizetimecomplexbuildindex}, showing the construction
-  times for the state machine and inverted index relative to the
-  average document size. The comparison shows that state machine
-  construction time is several orders of magnitude smaller than index
-  construction and negligible by comparison. Our state machine is
-  built in 770 ms on average for 1000 keywords with a standard
-  deviation of 18ms.
-
-  \begin{figure}[h!]
-    \begin{center}
-      \includegraphics[width=\textwidth,height=!]{docsizetimecomplexbuildsm}
-    \end{center}
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{docsizetimecomplexbuildsm}
     \caption{Time to build Aho-Corasick state machine as a function of
@@
     \label{fig:docsizetimecomplexbuildsm}
+  \end{minipage}
+  \hspace{0.5cm}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{docsizetimecomplexbuildindex}
+    \caption{Time to build an inverted index as a function of the
+      average document size. Time taken is shown in milliseconds on the
+      y-axis and the average document size in characters on the x-axis} 
+    \label{fig:docsizetimecomplexbuildindex}
+  \end{minipage}
 \end{figure}
 
-\item \textbf{Index Construction:} 
+State machine construction time does not depend on the
+length of the documents. Nonetheless it is interesting to compare
+figures \ref{fig:docsizetimecomplexbuildsm} and
+\ref{fig:docsizetimecomplexbuildindex}, showing the construction times
+for the state machine and inverted index relative to the average
+document size. The comparison shows that state machine construction
+time is several orders of magnitude smaller than index construction
+and negligible by comparison. Our state machine is built in 770 ms on
+average for 1000 keywords with a standard deviation of 18ms.
+
 As expected the time to build the indexer is almost linear with
@@
 
-  \begin{figure}[h!]
-    \begin{center}
-      \includegraphics[width=\textwidth,height=!]{docsizetimecomplexbuildindex}
-    \end{center}
-    \caption{Time to build an inverted index as a function of the
-      average document size. Time taken is shown in milliseconds on the
-      y-axis and the average document size in characters on the x-axis} 
-    \label{fig:docsizetimecomplexbuildindex}
-  \end{figure}
-
-\end{itemize}
-
-\subsubsection{Space Complexity}
-Here we are studying the number of records in our index database as a
-function of the average document size. As expected, the larger the documents
-the more records created to represent the index. However, this is not linear
-since if a word occurs many times in the same post we just save it
-once in the database. This relationship is shown in figure
+As expected, the larger the documents the more records created to
+represent the index. However, the relationship is not linear since if
+a word occurs many times in the same post we just save it once in the
+database. This relationship is shown in figure
 \ref{fig:docsizespacecomplexbuildindex}.
 
-\begin{figure}[h!]
+\begin{wrapfigure}{r}{0.70\textwidth}
   \begin{center}
-    \includegraphics[width=\textwidth,height=!]{docsizespacecomplexbuildindex}
+        \includegraphics[width=0.70\textwidth,height=!]{docsizespacecomplexbuildindex}
   \end{center}
@@
   \label{fig:docsizespacecomplexbuildindex}
-\end{figure}
+\end{wrapfigure}
 
@@
 
-\subsubsection{Complexity Analysis}
 For the na\"{i}ve algorithm described above, the theoretical worst
@@
 the state machine algorithm the theoretical complexity relationship is
-given by 
-equation \ref{eqn:timecomplexsm}.
+given by equation \ref{eqn:timecomplexsm}.
 
@@
 
-\subsubsection{Document Size}
-For this experiment we fixed the number of keywords to 1,000 and the
-number of documents to 5,000 and we ran our code on a database
-composed of document of different length. The experiment results are
-shown in figure \ref{fig:naivedocumentsize}.
-
-\begin{figure}[h!]
-  \begin{center}
-    \includegraphics[width=\textwidth,height=!]{naivedocumentsize}
-  \end{center}
-  \caption{Time to build an inverted index as a function of the
-      average document size. Time taken is shown in miliseconds on the
-      y-axis and the average document size in characters on the x-axis}
-  \label{fig:naivedocumentsize}
-\end{figure}
-
-As expected, the curves for both our indexer and the naive indexer are
+To exmaine the effect of document size, the number of keywords was
+fixed at 1,000 and the number of documents to 5,000 and we ran our
+code on a database composed of document of different length. The
+experiment results are shown in figure \ref{fig:naivedocumentsize}. As
+expected, the curves for both our indexer and the naive indexer are
 almost linear since for the naive implementation:
@@
 experiment results are plotted on a logarithmic scale in figure
-\ref{fig:naivedocumentsizelog}.
+\ref{fig:naivedocumentsizelog}.As expected, the two implementations
+exhibit a linear response to the number of keywords used. In the
+na\"{i}ve implemementation:
 
-\begin{figure}[h!]
-  \begin{center}
-    \includegraphics[width=\textwidth,height=!]{naivedocumentsizelog}
-  \end{center}
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{naivedocumentsize}
+    \caption{Time to build an inverted index as a function of the
+      average document size. Time taken is shown in miliseconds on the
+      y-axis and the average document size in characters on the x-axis}
+    \label{fig:naivedocumentsize}
+  \end{minipage}
+  \hspace{0.5cm}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{naivedocumentsizelog}
     \caption{Time to build an inverted index as a function of the
@@
     \label{fig:naivedocumentsizelog}
+  \end{minipage}
 \end{figure}
-
-\subsubsection{Number of Keywords}
-As expected, the two implementations exhibit a linear response to the
-number of keywords used. In the na\"{i}ve implemementation:
 
@@
 
-\begin{figure}[p]
-  \begin{center}
-    \includegraphics[width=\textwidth,height=!]{naivenumkeywords}
-  \end{center}
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{naivenumkeywords}
     \caption{Time to build an inverted index as a function of the
@@
     \label{fig:naivenumkeywords}
-\end{figure}
-
-\begin{figure}[p]
-  \begin{center}
-    \includegraphics[width=\textwidth,height=!]{naivenumkeywordslog}
-  \end{center}
+  \end{minipage}
+  \hspace{0.5cm}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{naivenumkeywordslog}
     \caption{Time to build an inverted index as a function of the
@@
     \label{fig:naivenumkeywordslog}
+  \end{minipage}
 \end{figure}
 
-\subsubsection{Size of Corpus}
 As expected, the time complexity is linearly related to the number of
-documents being indexed. The experiment results are shown in figure
-\ref{fig:naivesizecorpus} and on a logarithmic scale in figure
-\ref{fig:naivesizecorpuslog}. For the na\"{i}ve implementation:
+documents being indexed. The experiment results, in this respect, are
+shown in figure \ref{fig:naivesizecorpus} and on a logarithmic scale
+in figure \ref{fig:naivesizecorpuslog}. For the na\"{i}ve
+implementation:
 
@@
 
-\begin{figure}[p]
-  \begin{center}
-    \includegraphics[width=\textwidth,height=!]{naivesizecorpus}
-  \end{center}
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{naivesizecorpus}
     \caption{Time to build an inverted index as a function of the
@@
     \label{fig:naivesizecorpus}
-\end{figure}
-
-\begin{figure}[p]
-  \begin{center}
-    \includegraphics[width=\textwidth,height=!]{naivesizecorpuslog}
-  \end{center}
+  \end{minipage}
+  \hspace{0.5cm}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{naivesizecorpuslog}
     \caption{Time to build an inverted index as a function of the
@@
   \label{fig:naivesizecorpuslog}
+  \end{minipage}
 \end{figure}
@@
 investigate the effect of parallelization on the indexer operation, the
-indexer was deployed as follows:
-
-\begin{itemize}
-\item Single database server storing documents, keywords, and the
-  resulting index.
-\item Four indexer nodes each running a single instance of the indexer
-  program. 
-\end{itemize}
-
-This deployment as well as a description of the sequence of operation
-are shown in figure \ref{fig:paralleldeploymentmodel}.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[width=\textwidth,height=!]{paralleldeploymentmodel}
-  \end{center}
-  \caption{Deployment model of parallel indexer} 
-  \label{fig:paralleldeploymentmodel}
-\end{figure}
+indexer was deployed using a single database server storing documents,
+keywords, and the resulting index; four indexer nodes each running a
+single instance of the indexer program.
 
@@
 effect of corpus size on performance. The results of this experiment
-are shown in figure \ref{fig:parallelization} and described in detail
-in the following section.
+are shown in figure \ref{fig:parallelization}.
 
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[width=\textwidth,height=!]{parallelization}
-  \end{center}
-  \caption{Performance characteristics of parallel indexer
-    deployment. The \textit{ideal} curve shows the theoretical
-    performance curve based on linear combination of average
-    individual node performance over the corpus} 
-  \label{fig:parallelization}
-\end{figure}
-
-\subsubsection{Results}
 For the 100,000 document corpus, increasing the number of nodes
@@
 collection, this degradation is ultimately unavoidable.
+
+\begin{wrapfigure}{r}{0.70\textwidth}
+  \begin{center}
+        \includegraphics[width=0.60\textwidth,height=!]{parallelization}
+  \end{center}
+  \caption{Performance characteristics of parallel indexer
+    deployment. The \textit{ideal} curve shows the theoretical
+    performance curve based on linear combination of average
+    individual node performance over the corpus} 
+\end{wrapfigure}
 

### 2011-10-26 report/article.tex
Paper down to 11 pages - sending for review
Naive comparison almost done

@@
 scanning tool employed in the construction of an inverted
-index. Consistent with analytical predictions, an empirical evaluation
-of our hybrid indexer, shows significant improvement over a
-na\"{\i} implementation. In addition, the modular architecture of our indexer
-allowed the indexing task to be parallelized over multiple physical
-nodes. The results of parallelization are presented and offer
-interesting insights towards further horizontal scaling.
+index. Consistent with our analysis, an empirical evaluation of our
+indexer implementation shows significant improvement over a na\"{\i}ve
+algorithm. The modular architecture of our indexer allows the indexing
+task to be parallelized over multiple physical nodes. The results of
+parallelization are presented and offer interesting insights towards
+further horizontal scaling.
 \\\\
@@
 determine term frequencies. The Aho-Corasick string matching
-algorithm\cite{RefWorks:103} is a simple and efficient text scanning algorithm. The
-algorithm constructs a finite state machine to scan for a given set of
-keywords. It is, in effect, a reduced grammar regular expression
-parser of the type described in \cite{RefWorks:111}. The algorithm is simple and
-efficient, construction time being proportional to the sum of the
-length of the keyword set, and the number of state transitions
-required to scan a document is independent of the number of the size
-of the keyword set. Aho-Corasick was originally proposed as a
-bibliographic search mechanism, efficient enough to preclude the
-construction and maintenance of a search index. As the size of
-document collections have grown since the publishing of the
-Aho-Corasick algorithm, document indexing is essential for efficient
-document search. In this paper we demonstrate the use of Aho-Corasick,
-not as an alternative to indexing but as a lexical scanning tool
-employed in the construction of an inverted index. We provide analytic
-predictions of our implementation against a naive algorithm and
-finally, an empirical evaluation of our inverted index implementation
-is conducted and the results analyzed.
+algorithm\cite{RefWorks:103} is a simple and efficient text scanning
+algorithm. The algorithm constructs a finite state machine to scan for
+a given set of keywords. It is, in effect, a reduced grammar regular
+expression parser of the type described in \cite{RefWorks:111}. The
+algorithm is simple and efficient, construction time being
+proportional to the sum of the length of the keyword set, and the
+number of state transitions required to scan a document is independent
+of the number of the size of the keyword set. In this paper we
+demonstrate the use of Aho-Corasick, not as an alternative to indexing
+but as a lexical scanning tool employed in the construction of an
+inverted index. We provide analytic predictions of our implementation
+against a na\"{\i}ve algorithm and finally, an empirical evaluation of
+our inverted indexer implementation is conducted and the results
+analyzed.
 
@@
 \label{sec:implementation}
-As can be seen in figure \ref{fig:deploymentmodel}, the indexer
-processes are symmetrical and can deployed on multiple physical nodes. The
-individual indexers do not interact directly with one another, making
-for a simple deployment and operation model. The execution loop of
-each indexer is as follows:
+The Aho-Corasick indexer was implemented using Java and MySQL
+relational database. Upon startup, the indexer reads a list of index
+keywords and synonyms from the database and constructs an Aho-Corasick
+state machine. The indexer then enters a loop in which it uses the
+state machine to scan batches of documents retrieved from the
+database. The document batch size is varied manually to match the
+physical capabilites of the host. A possible enhancement to our
+implementation could include an adaptive loading component. The output
+of the state machine is used to augment a global index maintained in
+the database.
 
-\begin{enumerate}
-\item Initialize the indexer by retrieving a collection of index
-  keywords and synonyms from a relational database. These keywords are
-  used to construct a lexical parser which will be used to scan and
-  index documents.
-\item Retrieve a batch of unprocessed documents from a relational
-  database. The document batch will be sized according to the physical
-  capabilities of each node. In this implementation, this tuning task
-  is a manual exercise but future enhancements may include an adaptive
-  loading component.
-\item Parse each document retrieved and construct an inverted index
-  representing the batch. This \textit{delta} index, as we shall call
-  it, is then used to augment the global index maintained in a
-  relational database.
-\item Repeat from step 2.
-\end{enumerate}
-
-As shown in figure \ref{fig:deploymentmodel}, the searcher component
-can be executed from any node which has access to the relational
-database housing the document index. The execution path of a single
-search query wold be as follows:
-
-\begin{enumerate}
-  \item Canonize search terms based on keyword synonyms defined in the
-    keyword store.
-  \item Execute a boolean query against the inverted index.
-  \item Return the list of corpus documents containing the
-    intersection of the canons of the search terms.
-\end{enumerate}
+Figure \ref{fig:deploymentmodel} shows how multiple indexer instances
+can be deployed in parallel, using the database to synchronize their
+access to the corpus and index. The individual indexers do not
+interact directly with one another, making for a simple deployment and
+operation model.
 
@@
   \end{center}
-  \caption{Indexer deployment model}
+  \caption{Indexer deployment model showing multiple indexer instances
+  operating in parallel}
   \label{fig:deploymentmodel}
 \end{wrapfigure}
+
+Section \label{sec:experimentresults} describes the tests conducted to
+examine the effect of dictionary, document, and corpus size on the
+performance of a single Aho-Corasick indexer and in \label{sec:naive}
+we compare our indexer performance to that of a na\"{\i}ve
+implementation. Construction time of the state machine as a function
+of dictionary size was tested in order to quantify the overhead it
+imposes on the indexing task. In section \ref{parallelization} we
+explore the performance characteristics of a parallel indexer
+deployment.
+
+All tests were run on a single Intel Core 2 Duo 2GHz processor, 1GB
+RAM, running a 64-bit Linux 2.6.35 SMP kernel. The Java Virtual
+Machine used was version 1.6.0-24.
 
@@
 \label{sec:experimentresults}
-The primary goal of this project is to be able to reduce the time
-to build a document index. As our system is building the inverted index by
-parsing all documents of the collection, we can expect the
-time complexity of the task to be proportional to the sum total of
-characters in the collection. As a result, changing the  
-number of documents or the average size of each document should
-have a big impact in our time complexity. In addition, our
-implementation will incur the cost of building the Aho-Corasick state
-machines used to scan the documents. We can expect the time to build a
-state machine to be proportional to the sum of the size of the
-keywords.
-
-Finally we will do a time comparison between our indexer
-implementation and a na\"{i}ve implementation. All the time
-complexities tests are done on a single computer and it is important
-to understand that all these processes can be parallelized. In section
-\ref{parallelization} we install our indexer on multiple computers and
-explore the performance gain of running parallel instances
-simultaneously.
-
-All tests were run on a single Intel Core 2 Duo 2GHz processor, 1GB
-RAM, running a 64-bit Linux 2.6.35 SMP kernel. The Java Virtual
-Machine used was version 1.6.0-24. The database used was MySQL,
-running on the same node as the indexer.
 
@@
 \subsection{Number of Keywords}
-To be able to study the impact of different number of keywords, we
-built a keywords database with 10,000 most popular English
-words \cite{wordlist}. All the experiments in this
-part are done with 50,000 posts randomly downloaded from the Internet. The
-idea is to get a good distribution of documents with respect to length
-and content. The experiment shown here consist of building the index
-for different numbers of keywords from the 10,000 keyword database and
-explore the time and space complexity for the building of the state
-machine and the building of the index using the state machine.
-
-The time taken to construct the Aho-Corasick state machine was
-measured for different numbers of keywords. The results of this test
-is shown in figure \ref{fig:numkeywordstimecomplexbuildsm} and
-construction time can easily be considered linear with respect to the
-number of keywords. This is consistent with \cite{RefWorks:103} which
-proves that the state machine construction algorithm is linearly
-proportional to the sum of the lengths of the keywords used to
-construct the state machine.
-
-The time taken for a state machines to build an inverted index of our
-corpus was measured. The test was repeated using state machines
-constructed with a varying number of keywords and the results plotted
-in figure \ref{fig:numkeywordstimecomplexbuildindex}. According to
-\cite{RefWorks:103} which proves that the number of state transitions
-involved in processing an input string is independent of the number of
-keywords used to construct the state machine, we would not expect the
-time complexity to increase with the size of the keyword set. However,
-as can be seen in \ref{fig:numkeywordstimecomplexbuildindex}, the time
-to build the index increases logarithmically with the number of
-keywords. This is likely due to the fact that as the number of
-keywords increases, so does the size of the index and the number of
-database interacts required to store and update the index.
+To study the impact of dictionary size on performance keywords, a
+10,000 keyword dictionary of the most popular English words
+\cite{wordlist} and a corpus of 50,000 blog posts of varying 
+size were used. First, the time taken to construct the Aho-Corasick state
+machine was measured for different numbers of dictionary keywords. The
+results of this test, shown in figure
+\ref{fig:numkeywordstimecomplexbuildsm}, indicate construction time to be
+linear with respect to the number of keywords. This is consistent with
+\cite{RefWorks:103} which formally proves that the state machine
+construction algorithm is linearly proportional to the sum of the
+lengths of the keywords used to construct the state machine.
 
@@
 
-The space used by the index in the database is proportional to the
-number of words we are using. It also means that our words have the
-same probability at the beginning or at the end of our 12,000 word
-dictionary.
+Second, the time taken for a state machines to build an inverted index
+of our corpus was measured. The test was repeated using state machines
+constructed with a varying number of keywords and the results plotted
+in figure \ref{fig:numkeywordstimecomplexbuildindex}. According to
+\cite{RefWorks:103}, which proves that the number of state transitions
+involved in processing an input string is independent of the number of
+keywords used to construct the state machine, we would not expect the
+time complexity to increase with the size of the keyword set. However,
+as can be seen in \ref{fig:numkeywordstimecomplexbuildindex}, the time
+to build the index increases logarithmically with the number of
+keywords. This is likely due to the fact that as the number of
+keywords increases, so does the size of the index and the amount of
+database interactions required to update the index.
+
+It is important to note that for this experiment, the time taken to
+construct the state machine is two orders of magnitude lower than the
+time taken by the indexing task. This is an important result as it
+validates the practical use of the Aho-Corasick algorithm to index
+large corpora, the overhead of state machine construction time being
+negligible relative to the indexing task. Also, the space used by the
+index in the database is proportional to the number of words we are
+using.
 
@@
 \subsection{Size of Corpus}
-To be able to study the impact of different number of Posts, we built
-a posts database with 400,000 posts randomly downloaded from
-the Internet. Then the keywords used to build the index are simply the
-first 5,000 from the previous keywords database.
 
-The experiment shown here consist on building the index for different
-number of Posts and explore the time and space complexity for the
-building of the state machine and the building of the index once the
-state machine is built. We can expect that the time to build the State
-machine won’t change with the number of posts and the time to build
-the index will increase almost linearly with the number of posts.
-
-As we were expecting, the time to build the index with the state
-machine strategy is linear with respect to the number of post, as
-depicted in figure \ref{fig:corpsizetimecomplexbuildindex}. However, to be able
-to perform this experiment we had to run the indexer program with
-batches of 10,000 posts due to Java heap space problems. As a result, the state
-machine is rebuilt every 10,000 posts.
-
-The space used to store the index in the database is
-proportional to the size of the corpus as shown in figure
-\ref{fig:corpsizespacecomplexbuildindex}.
-
-\begin{figure}[ht]
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
-    \includegraphics[width=\textwidth]{corpsizetimecomplexbuildindex}
+\begin{wrapfigure}{r}{0.70\textwidth}
+  \begin{center}
+        \includegraphics[width=0.70\textwidth,height=!]{corpsizetimecomplexbuildindex}
+  \end{center}
     \caption{Time to build the inverted index as a function of the
@@
     \label{fig:corpsizetimecomplexbuildindex}
-  \end{minipage}
-  \hspace{0.5cm}
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
-    \includegraphics[width=\textwidth]{corpsizespacecomplexbuildindex}
-    \caption{Size of index as a function of the size of the
-      corpus. The number of database rows used to store the index is
-      shown on the y-axis and the number of documents in the
-      collection on the x-axis} 
-    \label{fig:corpsizespacecomplexbuildindex}
-  \end{minipage}
-\end{figure}
+\end{wrapfigure}
+
+To examine the effect of corpus size on the performance of the
+indexer, a 5,000 keyword dictionary and a corpus of 800,000 blog posts
+of various sizes were used. The timing results of indexing corpora of
+varying sizes, shown in figure
+\ref{fig:corpsizetimecomplexbuildindex}, indicate that the time to
+build the index with the state machine is linear with respect to the
+number of posts.
 
@@
 \subsection{Document Size} 
-As we have seen before, the time to build the index is almost
-proportional with the number of post. Theoretically this time should
-be proportional to the total length of all posts. Let’s focus on this
-aspect and study the impact of the length of the post on the time to
-build the state machine. To do that we built a 5,000 posts database
-with posts that are all composed of 20,000 characters. Then, we ran
-our system on different length of substring of these posts.
+To examine the effect of document size on indexer performance, 
+a 5,000 keyword dictionary and 5,000 blog posts were used. Tests were
+run in which the the size of the blog posts was varied, generating the
+performance curve in figure \ref{fig:docsizetimecomplexbuildindex}.  
+As expected the time to build the indexer is almost linear with
+respect to the average document size.
 
 \begin{figure}[ht]
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
-    \includegraphics[width=\textwidth]{docsizetimecomplexbuildsm}
-    \caption{Time to build Aho-Corasick state machine as a function of
-      the average document size. Time taken is shown in milliseconds on the
-      y-axis and the average document size in characters on the x-axis}
-    \label{fig:docsizetimecomplexbuildsm}
-  \end{minipage}
-  \hspace{0.5cm}
   \begin{minipage}[b]{0.5\linewidth}
@@
   \end{minipage}
-\end{figure}
-
-State machine construction time does not depend on the
-length of the documents. Nonetheless it is interesting to compare
-figures \ref{fig:docsizetimecomplexbuildsm} and
-\ref{fig:docsizetimecomplexbuildindex}, showing the construction times
-for the state machine and inverted index relative to the average
-document size. The comparison shows that state machine construction
-time is several orders of magnitude smaller than index construction
-and negligible by comparison. Our state machine is built in 770 ms on
-average for 1000 keywords with a standard deviation of 18ms.
-
-As expected the time to build the indexer is almost linear with
-respect to the average document size. This relationship is shown in
-figure \ref{fig:docsizetimecomplexbuildindex}.
-
-As expected, the larger the documents the more records created to
-represent the index. However, the relationship is not linear since if
-a word occurs many times in the same post we just save it once in the
-database. This relationship is shown in figure
-\ref{fig:docsizespacecomplexbuildindex}.
-
-\begin{wrapfigure}{r}{0.70\textwidth}
-  \begin{center}
-        \includegraphics[width=0.70\textwidth,height=!]{docsizespacecomplexbuildindex}
-  \end{center}
+  \hspace{0.5cm}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=\textwidth]{docsizespacecomplexbuildindex}
     \caption{Size of index as a function of the average document
@@
     \label{fig:docsizespacecomplexbuildindex}
-\end{wrapfigure}
+  \end{minipage}
+\end{figure}
+
+Examining the size of the index created, the larger the documents the
+more records created. However, the relationship is not linear since a
+single record is stored to represent one or many occurrences of a
+particular keyword in a document. This relationship is shown in figure  
+\ref{fig:docsizespacecomplexbuildindex}.
 
@@
 \subsection{Comparison with Na\"{i}ve Index Builder}
-This part focuses on the improvement that our system with respect to
-time complexity in comparison with a na\"{i}ve index builder. The na\"{i}ve
-indexer builder has the same input as our system i.e. a list of
-keywords and a list of documents, and the same output i.e. an inverted
-index corresponding to the keywords and documents. Algorithm
-\ref{alg:invertedindex} is used in principal for both implementations.
+\label{sec:naive}
+In order to quantify the benefits of using the Aho-Corasick algorithm
+for indexing, our indexer performace was compared to that of a
+na\"{i}ve implementation. The na\"{i}ve indexer builder has the same
+input as our system i.e. a list of keywords and a list of documents,
+and the same output i.e. an inverted index corresponding to the
+keywords and documents. Algorithm \ref{alg:invertedindex} is used in
+principal for both implementations. However, whereas the Aho-Corasick
+implementation uses a state machine to scan, the na\"{\i}ve
+indexer tokenizes the document and compares each token to the keyword
+dictionary.
 
@@
 
-As we can see this algorithm has a complexity that directly depends on
-the number of keywords $n$, the number of document $m$, and the length
-of document since we have to check in each document for each
-keyword. Moreover, we decided to limit, as much as possible, the
-accesses to the database since it is really time consuming in
-Java. That’s why we do one access to the database for each keywords
-and not for each document. In the state machine we do as much accesses
-to the database as the number of document (in the worst case meaning
-we find at least one keyword per document). In the following
-subsections we compare the results for this algorithm with our system
-results for different number of keywords and posts.
+Algorithm \ref{alg:invertedindex} has a complexity proportional to the
+number of keywords $n$, the number of document $m$, and the average
+document size $\bar{c}$. For our chosen architecture, execution time
+is also effected by database latency. In the case of the
+na\"{\i}ve indexer, we perform one database transaction per keyword
+and in the Aho-Corasick implementation, we perform a database
+transaction per document. For analysis purposes, we designate the
+average database latency per transaction as $\bar{t}$.
 
-For the na\"{i}ve algorithm described above, the theoretical worst
-case time complexity relationship is shown in equation
-\ref{eqn:timecomplexworstcase}.
+For the na\"{i}ve indexer implementation described above, the expected
+time complexity relationship is shown in equation
+\ref{eqn:timecomplexworstcase}. For the Aho-Corasick indexer, the
+expected theoretical complexity relationship is given by equation
+\ref{eqn:timecomplexsm}.
 
@@
 \label{eqn:timecomplexworstcase}
-Time \propto keyNum * ( docNum * maxLength + conn)
+Time \propto n * ( m * \bar{c} + \bar{t} )
 \end{equation}
-
-where \(keyNum\) is the number of keywords, \(docNum\) is the number
-of documents, \(maxLength\) is the maximum document length, and
-\(conn\) is the time taken to create a connection to the database. For
-the state machine algorithm the theoretical complexity relationship is
-given by equation \ref{eqn:timecomplexsm}.
 
@@
 \label{eqn:timecomplexsm}
-Time \propto docNum * ( maxLength + conn)
+Time \propto m * ( \bar{c}  + \bar{t} )
 \end{equation}
 
-To exmaine the effect of document size, the number of keywords was
-fixed at 1,000 and the number of documents to 5,000 and we ran our
-code on a database composed of document of different length. The
-experiment results are shown in figure \ref{fig:naivedocumentsize}. As
-expected, the curves for both our indexer and the naive indexer are
-almost linear since for the naive implementation:
-
-\[Time \propto (keyNum * docNum) * maxLength + conn * keyNum\]
-
-and for our Aho-Corasick state machine method:
-
-\[Time \propto docNum * maxLength + conn * docNum\]
-
-We can observe that for a small length of document naive method is
-faster since \(keyNum\) is smaller than \(docNum\). For \(docLength = 1300\)
-our indexer is far more efficient. To see this effect more clearly,
-experiment results are plotted on a logarithmic scale in figure
-\ref{fig:naivedocumentsizelog}.As expected, the two implementations
-exhibit a linear response to the number of keywords used. In the
-na\"{i}ve implemementation:
+To examine the effect of document size on idexer performance, the number of keywords was
+fixed at 1,000, the number of documents at 5,000, and tests run
+against corpora composed of varying average length. From equations
+\ref{eqn:timecomplexworstcase} and \ref{eqn:timecomplexsm}, we would
+expect both implementations to behave linearly with repect to average
+documetn size and to differ by constant scaling factor $n$, the size
+of the keyword dictionary. The experiment results are shown in figure
+\ref{fig:naivedocumentsize} and support this analysis. Also, we can observe
+that for smaller documents, the na\"{\i}ve method
+performs better. This is attributable to the fact that for smaller
+documents, the database advantage of the na\"{\i}ve implementation is
+enough to overcome the keyword dictionary size scaling factor. To see
+this effect more clearly, experiment results are plotted on a
+logarithmic scale in figure \ref{fig:naivedocumentsizelog}.
 
@@
 
-\[Time \propto (keyNum * maxLength) * docNum + conn * keyNum\]
-
-and in our Aho-Corasick state machine method:
-
-\[Time \propto (maxLength + conn) * docNum \]
+xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
 
@@
     individual node performance over the corpus} 
+  \label{fig:parallelization}
 \end{wrapfigure}

### 2012-02-17 report/article.tex

@@
 database. The document batch size is varied manually to match the
-physical capabilites of the host. A possible enhancement to our
+physical capabilities of the host. A possible enhancement to our
 implementation could include an adaptive loading component. The output
@@
 %----------------------------------------
-% Comparison with Naive Index Builder
+% Comparison with Naive Indexer
 %----------------------------------------
-\subsection{Comparison with Na\"{i}ve Index Builder}
+\subsection{Comparison with Na\"{i}ve Indexer}
 \label{sec:naive}
 In order to quantify the benefits of using the Aho-Corasick algorithm
-for indexing, our indexer performace was compared to that of a
+for indexing, our indexer performance was compared to that of a
 na\"{i}ve implementation. The na\"{i}ve indexer builder has the same
@@
 Algorithm \ref{alg:invertedindex} has a complexity proportional to the
-number of keywords $n$, the number of document $m$, and the average
+number of keywords $n$, the number of documents $m$, and the average
 document size $\bar{c}$. For our chosen architecture, execution time
@@
 
-To examine the effect of document size on idexer performance, the number of keywords was
-fixed at 1,000, the number of documents at 5,000, and tests run
-against corpora composed of varying average length. From equations
-\ref{eqn:timecomplexworstcase} and \ref{eqn:timecomplexsm}, we would
-expect both implementations to behave linearly with repect to average
-documetn size and to differ by constant scaling factor $n$, the size
-of the keyword dictionary. The experiment results are shown in figure
-\ref{fig:naivedocumentsize} and support this analysis. Also, we can observe
-that for smaller documents, the na\"{\i}ve method
-performs better. This is attributable to the fact that for smaller
-documents, the database advantage of the na\"{\i}ve implementation is
-enough to overcome the keyword dictionary size scaling factor. To see
-this effect more clearly, experiment results are plotted on a
-logarithmic scale in figure \ref{fig:naivedocumentsizelog}.
+In the context of these equations, we describe and interpret our
+empirical results in the following subsections, testing for the effects
+of document size $c$, number of keywords $n$, and number of documents
+$m$.
+
+%----------------------------------------
+% Document Size
+%----------------------------------------
+\subsubsection{Document Size}
+To examine the effect of document size on indexer performance, the
+number of keywords was fixed at 1,000, the number of documents at
+5,000, and tests run against corpora composed of varying average
+length. From equations \ref{eqn:timecomplexworstcase} and
+\ref{eqn:timecomplexsm}, we would expect both implementations to
+behave linearly with respect to average document size and to differ by
+a constant scaling factor $n$, the size of the keyword dictionary. The
+experiment results are shown in figure \ref{fig:naivedocumentsize} and
+support this analysis. Also, we observe that for smaller
+documents, the na\"{\i}ve method performs better. This is attributable
+to the fact that for smaller documents, the database advantage of the
+na\"{\i}ve implementation is enough to overcome the keyword dictionary
+size scaling factor. To see this effect more clearly, experiment
+results are plotted on a logarithmic scale in figure
+\ref{fig:naivedocumentsizelog}.
 
@@
     \caption{Time to build an inverted index as a function of the
-      average document size. Time taken is shown in miliseconds on the
+      average document size. Time taken is shown in milliseconds on the
       y-axis and the average document size in characters on the x-axis}
@@
 
-xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
-
-However, we would have expected the state machine method not to depend
-on the number of keyword. The reason why it’s not the case is because
-when we don’t have any keywords in a document, we don’t access the
-database. Knowing that the database access is very time consuming,
-increasing the number of keyword increases the probability of finding a
-keyword into each document and so increase the time complexity.
+%----------------------------------------
+% Keywords
+%----------------------------------------
+\subsubsection{Number of Keywords}
+To examine the effect of keyword dictionary size on indexer
+performance, the number of documents was fixed at 5,000, and tests run
+using varying numbers of keywords. From equation
+\ref{eqn:timecomplexworstcase}, we would expect the na\"{\i}ve
+implementation to have a linear dependency on the number of keywords
+and this is supported by the test results. From equation
+\ref{eqn:timecomplexsm} we would not expect the Aho-Corasick
+implementation to depend on the number of keyword however, the test
+results seem to indicate a linear dependency. This can be attributed
+to the fact that the larger the keyword dictionary, the more likely we
+are to find matches in a document and incur the expensive database
+latency cost.
 
@@
 
-As expected, the time complexity is linearly related to the number of
-documents being indexed. The experiment results, in this respect, are
-shown in figure \ref{fig:naivesizecorpus} and on a logarithmic scale
-in figure \ref{fig:naivesizecorpuslog}. For the na\"{i}ve
-implementation:
-
-\[Time \propto (docNum * maxLenght + conn) * keyNum\]
-
-and in our Aho-Corasick state machine method:
-
-\[Time \propto docNum * maxLenght + conn * docNum\]
+%----------------------------------------
+% Corpus Size
+%----------------------------------------
+\subsubsection{Number of Documents}
+From equations \ref{eqn:timecomplexworstcase} and
+\ref{eqn:timecomplexsm}, the time complexity is linearly related to
+the number of documents being indexed. The experiment results, in this
+respect, are shown in figure \ref{fig:naivesizecorpus} and on a
+logarithmic scale in figure \ref{fig:naivesizecorpuslog}.
 
@@
 indexer was deployed using a single database server storing documents,
-keywords, and the resulting index; four indexer nodes each running a
-single instance of the indexer program.
+keywords, and the resulting index. Four indexer nodes were used, each
+running a single instance of the indexer program.
 
@@
 normalized to a throughput metric (documents/second) for convenient
-comparison. Test results using the same number of nodes were averaged
-to account for differences in the physical capabilities and
-configuration of the worker nodes. Next, a series of tests was
+comparison and to account for differences in the physical capabilities
+and configuration of the worker nodes. A series of tests were
 performed using a corpus of 800,000 documents to investigate the
@@
 are shown in figure \ref{fig:parallelization}.
+
+\begin{wrapfigure}{r}{0.70\textwidth}
+  \begin{center}
+        \includegraphics[width=0.60\textwidth,height=!]{parallelization}
+  \end{center}
+  \caption{Performance characteristics of parallel indexer
+    deployment. The \textit{ideal} curve shows the theoretical
+    performance curve based on linear combination of average
+    individual node performance over the corpus} 
+  \label{fig:parallelization}
+\end{wrapfigure}
 
@@
 collection, this degradation is ultimately unavoidable.
-
-\begin{wrapfigure}{r}{0.70\textwidth}
-  \begin{center}
-        \includegraphics[width=0.60\textwidth,height=!]{parallelization}
-  \end{center}
-  \caption{Performance characteristics of parallel indexer
-    deployment. The \textit{ideal} curve shows the theoretical
-    performance curve based on linear combination of average
-    individual node performance over the corpus} 
-  \label{fig:parallelization}
-\end{wrapfigure}
 
